### 2022-01-28 Descriptives.Rmd
Created descriptives file, and added initial data loading

---
title: "Descriptives"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data
```{r}
struc_include<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title")

sum(struc_include$imgincl_t1w_include)
sum(struc_include$imgincl_t2w_include)
```

Data cleaning
```{r}

```