### 2022-02-01 Descriptives.Rmd
Added descriptives and plots

---
title: "Descriptives"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "reshape2",
               "nlme", "lme4",
               "data.table", "psych",
               "parallel","lubridate",
               "ggpubr", "broom", 
               "apaTables", "MetBrewer")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)
```

Load data
```{r}
struc_include<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:12) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include),
        imgincl_t2w_include = as.numeric(imgincl_t2w_include))

struc_include_baseline <- struc_include %>% 
  filter(eventname == "baseline_year_1_arm_1")

struc_include_followup <- struc_include %>% 
  filter(eventname == "2_year_follow_up_y_arm_1")

```

Basic descriptives
```{r}
t1_baseline = sum(struc_include_baseline$imgincl_t1w_include)
t2_baseline = sum(struc_include_baseline$imgincl_t2w_include)

t1_followup = sum(struc_include_followup$imgincl_t1w_include)
t2_followup = sum(struc_include_followup$imgincl_t2w_include)

t1_baseline_age_mean = mean(struc_include_baseline$interview_age[struc_include_baseline$imgincl_t1w_include==1], na.rm=T)
t2_baseline_age_mean = mean(struc_include_baseline$interview_age[struc_include_baseline$imgincl_t2w_include==1], na.rm=T)
t1_baseline_age_sd = sd(struc_include_baseline$interview_age[struc_include_baseline$imgincl_t1w_include==1], na.rm=T)

t1_followup_age_mean = mean(struc_include_followup$interview_age[struc_include_followup$imgincl_t1w_include==1], na.rm=T)
t2_followup_age_mean = mean(struc_include_followup$interview_age[struc_include_followup$imgincl_t2w_include==1], na.rm=T)
t1_followup_age_sd = sd(struc_include_followup$interview_age[struc_include_followup$imgincl_t1w_include==1], na.rm=T)

```

Descriptive plots
```{r}
ggplot(struc_include_baseline %>% filter(imgincl_t1w_include==1), aes(x=interview_age)) + 
  geom_histogram(binwidth=1, fill=met.brewer("Hokusai2", 1), color=met.brewer("Hokusai3", 1)) +
  theme_classic() +
  labs(x="Age (Months)", title = "Baseline Age Distribution")
ggsave("BaselineAgeDist.png")

ggplot(struc_include_followup %>% filter(imgincl_t1w_include==1), aes(x=interview_age)) + 
  geom_histogram(binwidth=1, fill=met.brewer("Hokusai2", 1), color=met.brewer("Hokusai3", 1)) +
  theme_classic() +
  labs(x="Age (Months)", title = "2yr Follow-up Age Distribution")
ggsave("2YrFollowupAgeDist.png")

```