### 2022-02-17 DataSetup.Rmd
Created Rmd for importing data

---
title: "DataSetup"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "reshape2",
               "nlme", "lme4",
               "data.table", "psych",
               "parallel","lubridate",
               "ggpubr", "broom", 
               "apaTables", "MetBrewer")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)
```

Load Structural Data
```{r}
#Load list of included subjects
struc_include<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:12) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include),
        imgincl_t2w_include = as.numeric(imgincl_t2w_include)) %>% 
  filter(imgincl_t1w_include == 1)

#Load structural data
#Filter for volume & area measurements
smri_1 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include$src_subject_id) %>% 
  select(matches("vol|area"))

#T1 intensity, might not need
smri_2 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip20201.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include$src_subject_id)

#Rename data to fit naming conventions
#????

```

Load Other Data
```{r}
#Puberty
youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
  filter(!collection_title=="collection_title")

youth_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphy01.txt") %>%
  filter(!collection_title=="collection_title")

parent_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
  filter(!collection_title=="collection_title")

parent_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphp01.txt") %>%
  filter(!collection_title=="collection_title")

saliva <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_hsss01.txt") %>%
  filter(!collection_title=="collection_title")

#Cognition
cognition <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, nihtbx_fluidcomp_fc, nihtbx_cryst_fc, nihtbx_totalcomp_fc)

#Height/weight metrics
physical <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ant01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, anthroheightcalc, anthroweightcalc)

#Sexual attraction and behavior
gish <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
  filter(!collection_title=="collection_title")

#Grade level, parent report
grade <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(subjectkey, src_subject_id, interview_date, interview_age, sex, eventname, demo_ed_v2)
```

```{r}

```