### 2022-02-17 ExisitingModel.Rmd

---
title: "Existing Model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Install XGboost & tidymodels
packageUrl <- "https://cran.r-project.org/src/contrib/Archive/xgboost/xgboost_1.0.0.2.tar.gz"
# You then install this version of the package using
install.packages(packageUrl, repos = NULL, type = 'source')

install.packages("tidymodels")

library(here) # OS agnostic relative paths (relative to project dir)
library(tidyverse) # data wrangling tools and pipes
library(tidymodels) # machine learning metaverse
library(xgboost) # main engine, needs version 1.0.0.2, see above

```

Load Data
```{r}

#load("smri_1.rds")

xgb_mod <- readRDS(
  file = here::here(
    "xgboost_9to19_brain_age_mod.rds"))
```

Predict Brain Age
```{r}
brain_age_df <-
  xgb_mod %>%
  predict(new_data = renamed_smri) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = renamed_smri$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)
```

Evaluate accuracy
```{r}
# compute common performance metrics: mae, rsq, rmse
brain_age_df %>%
  metrics(truth = truth, estimate = .pred)
```

Optional bias correction
```{r}
# bias prediction method, steps in comments,
# previously determined values hardcoded.
# xgb_bias_mod <- lm(.pred ~ truth, data = xgb_validate)

# extract intercept and slope
bias_intercept <- 6.41 # xgb_bias_mod$coefficients[["(Intercept)"]]
bias_slope <-  0.55 # xgb_bias_mod$coefficients[["truth"]]

# create bias correct data frame
brain_age_corrected_df <- brain_age_df %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - bias_intercept) / bias_slope
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

#mine
brain_age_corrected_df %>%
  metrics(truth = truth, estimate = corrected_pred)

```

plotting
```{r}
brain_age_actual_df <- as.data.frame(brain_age_df)

ggplot(brain_age_actual_df, aes(x=truth, y=.pred)) +
  geom_point()

ggplot(brain_age_actual_df, aes(x=.pred)) +
  geom_histogram()

ggplot(brain_age_actual_df, aes(x=truth)) +
  geom_histogram()

ggplot(as.data.frame(brain_age_corrected_df), aes(x=truth, y=.pred)) +
  geom_point()

ggplot(as.data.frame(brain_age_corrected_df), aes(x=corrected_pred)) +
  geom_histogram()
```