### 2022-03-18 Flux2022.Rmd
Created Rmd for Flux abstract - imports brainage, puberty, and cogntion. Cleans puberty scores, runs descriptives, splits into two sample, runs models, and tests model output.

---
title: "Flux2022"
author: "Lucy Whitmore"
date: "3/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(psych )
```

## Load Data
```{r, include=FALSE}
struc_include<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:12) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include),
        imgincl_t2w_include = as.numeric(imgincl_t2w_include)) %>% 
  filter(imgincl_t1w_include == 1)

#Load structural data
#Filter for volume & area measurements
smri_1 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include$src_subject_id) %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area"))

smri_baseline <- smri_1%>% 
  filter(eventname=="baseline_year_1_arm_1")

```

```{r}
#Puberty
#Raw scores, youth
youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  filter(eventname == "baseline_year_1_arm_1") %>% 
  select(5:38) %>% 
  select(-c(interview_date, pds_remote___1, pds_remote___2, pds_remote___3, pds_remote___4, pds_device)) %>% 
  mutate_at(c(5:28), as.numeric)

#Raw scores, parent
parent_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  filter(eventname== "baseline_year_1_arm_1") %>% 
  select(5:27) %>% 
  select(-c(interview_date, pds_select_language___1)) %>% 
  mutate_at(c(5:21), as.numeric)

#Cognition Scores (Age-Corrected)
cognition <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  filter(eventname == "baseline_year_1_arm_1") %>% 
  select(src_subject_id,interview_age, nihtbx_fluidcomp_agecorrected, nihtbx_cryst_agecorrected, nihtbx_totalcomp_agecorrected) %>% 
  mutate_at(c(2:5), as.numeric)

#Existing Models
corrected_baseline_brainage <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/BaselineBrainAgeCorrected.Rda")
```

```{r, eval=F}
#future data import (if we find data)

#Sum scores (categorical)
youth_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphy01.txt") %>%
  filter(!collection_title=="collection_title")

parent_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphp01.txt") %>%
  filter(!collection_title=="collection_title")

#Mean Scores & maybe all scores??
youth_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_mhy01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(src_subject_id, interview_age, sex, eventname,
         pds_y_ss_male_sum, pds_y_ss_male_sum_nm, pds_y_ss_male_mean, 
         pds_y_ss_male_mean_nm, pds_y_ss_male_category, pds_y_ss_male_cat_nm, 
         pds_y_ss_female_sum, pds_y_ss_female_sum_nm, pds_y_ss_female_mean, 
         pds_y_ss_female_mean_nm, pds_y_ss_female_category, pds_y_ss_female_cat_nm)

parent_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_mhp01.txt") %>%
  filter(!collection_title=="collection_title")

```

## Format BrainAge
```{r}
corrected_baseline_brainage$src_subject_id <- smri_baseline$src_subject_id
```
## Format Puberty Data 
```{r}
#two or more missing values = exclude
#PDS mean score is a continuous value calculated by averaging across PDS items following Herting et al. (2021). Items on the PDS were considered missing if participants answered, “I don’t know” or “refuse to answer”, if the response was left blank, or if the response value was outside of the expected range (1 to 4). Participants with two or more missing answers were excluded from the PDS score calculation (following recommendations from (Herting et al., 2021)

youth_pds[youth_pds == 777] <- NA
youth_pds[youth_pds == 999] <- NA

#Youth-report, removes everyone w/o complete data
youth_pds <- youth_pds %>% 
  mutate(youth_sum = case_when(pds_sex_y == "1" ~ #male
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_m4_y + pds_m5_y),  
                         pds_sex_y == "2" ~ #female
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_f4_2_y + pds_f5_y))) %>% 
  mutate(youth_mean = youth_sum/5)

youth_pds_use <- youth_pds %>% 
  select(src_subject_id, pds_sex_y, youth_sum, youth_mean)

#Parent-report, removes everyone w/o complete data
parent_pds[parent_pds == 777] <- NA
parent_pds[parent_pds == 999] <- NA
parent_pds <- parent_pds %>% 
    mutate(parent_sum = case_when(sex == "M" ~ #male
                           (pds_1_p + pds_2_p + pds_3_p + pds_m4_p + pds_m5_p),  
                         sex == "F" ~ #female
                           (pds_1_p + pds_2_p + pds_3_p + pds_f4_p + pds_f5b_p))) %>% 
  mutate(parent_mean = parent_sum/5)

parent_pds_use <- parent_pds %>% 
  select(src_subject_id, sex, parent_sum, parent_mean)
```

## Make big dataframe
```{r}
#Combine dataframes
brainage_pds <- corrected_baseline_brainage[, c(6,1:5)] %>% 
  left_join(youth_pds_use) %>% 
  left_join(parent_pds_use) %>% 
  left_join(cognition)

#Split sample
smp_size <- floor(0.5 * nrow(brainage_pds))

## set the seed to make your partition reproducible
set.seed(123)
sample_ind <- sample(seq_len(nrow(brainage_pds)), size = smp_size)

sample_1 <- brainage_pds[sample_ind, ]
sample_2 <- brainage_pds[-sample_ind, ]
```

## Plot Descriptives
```{r}
ggplot(brainage_pds, aes(x=interview_age)) + 
  geom_histogram()

ggplot(brainage_pds, aes(x=corrected_gap)) + 
  geom_histogram()

ggplot(brainage_pds, aes(x=youth_mean)) + 
  geom_histogram()

ggplot(brainage_pds, aes(x=parent_mean)) + 
  geom_histogram()

ggplot(brainage_pds, aes(x=nihtbx_totalcomp_agecorrected)) + 
  geom_histogram()

```

## Split Sample Descriptives
```{r}
#Sample 1 Descriptives
describe(sample_1)

#Sample 2 Descriptives
describe(sample_2)

```

## Run Models
```{r}
# brainage ~ youth sum
baseline_puberty_y_sum_1 <- lm(corrected_gap ~ youth_sum, data = sample_1)
baseline_puberty_y_sum_2 <- lm(corrected_gap ~ youth_sum, data = sample_2)

# brainage ~ youth mean
baseline_puberty_y_mean_1 <- lm(corrected_gap ~ youth_mean, data = sample_1)
baseline_puberty_y_mean_2 <- lm(corrected_gap ~ youth_mean, data = sample_2)

# brainage ~ parent sum
baseline_puberty_p_sum_1 <- lm(corrected_gap ~ parent_sum, data = sample_1)
baseline_puberty_p_sum_2 <- lm(corrected_gap ~ parent_sum, data = sample_2)

# brainage ~ parent mean
baseline_puberty_p_mean_1 <- lm(corrected_gap ~ parent_mean, data = sample_1)
baseline_puberty_p_mean_2 <- lm(corrected_gap ~ parent_mean, data = sample_2)

baseline_puberty_p_mean_flipped_1 <- lm(parent_mean ~ corrected_gap, data = sample_1)
baseline_puberty_p_mean_flipped_2 <- lm(parent_mean ~ corrected_gap, data = sample_2)

#Cognition
baseline_cognition_1 <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected, data=sample_1)
baseline_cognition_2 <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected, data=sample_2)

```

## Model Summaries
```{r}
#summary(baseline_puberty_y_sum_1)

summary(baseline_puberty_y_mean_1)
summary(baseline_puberty_y_mean_2)

#summary(baseline_puberty_p_sum)

summary(baseline_puberty_p_mean_1)
summary(baseline_puberty_p_mean_2)

summary(baseline_puberty_p_mean_flipped_1)
summary(baseline_puberty_p_mean_flipped_2)

summary(baseline_cognition_1)
summary(baseline_cognition_2)
```