### 2022-06-07 ABCDModelMultiwave.Rmd

---
title: "ABCDModelMultiwave"
author: "Lucy Whitmore"
date: "5/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "reshape2",
               "nlme", "lme4",
               "data.table", "psych",
               "parallel","lubridate",
               "ggpubr", "broom", 
               "apaTables", "MetBrewer")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)
```

Data Import/Set-Up
```{r}
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/abcd_train_multiwave.Rds")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/abcd_test_multiwave.Rds")

```

Test Model
```{r}
abcd_brainage_multiwave_model <- readRDS("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/xgboost_abcd_multiwave_brain_age_mod.rds")
```

Predict Brain Age (baseline)
```{r}
brain_age_df_multiwave <-
  abcd_brainage_multiwave_model %>%
  predict(new_data = abcd_test_multiwave) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = abcd_test_multiwave$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

```

Evaluate accuracy
```{r}
# compute common performance metrics: mae, rsq, rmse
brain_age_df_multiwave %>%
  metrics(truth = truth, estimate = .pred)
```

Optional bias correction
```{r}
# bias prediction method, steps in comments,
# previously determined values hardcoded.

#get validation set
set.seed(42)
df_split_multiwave <- initial_split(
  abcd_train_multiwave, 
  prop = 0.80,
  # matching age distributions across train and test set
  strata = "interview_age"
)

df_validation_multiwave <- testing(df_split_multiwave)

#think I need to now run model on validation set??
abcd_multiwave_brain_age_correction <-
  abcd_brainage_multiwave_model %>%
  predict(new_data = df_validation_multiwave) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = df_validation_multiwave$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

abcd_bias_mod_multiwave <- lm(.pred ~ truth, data = abcd_multiwave_brain_age_correction)

# extract intercept and slope
#abcd_bias_mod_multiwave$coefficients[["(Intercept)"]]
abcd_bias_intercept_multiwave <- 6.203475
#abcd_bias_mod_multiwave$coefficients[["truth"]]
abcd_bias_slope_multiwave <- 0.4300184

# create bias correct data frame
#Baseline
brain_age_corrected_df_multiwave <- brain_age_df_multiwave %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - abcd_bias_intercept_multiwave) / abcd_bias_slope_multiwave
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

#mine
brain_age_corrected_df_multiwave %>%
  metrics(truth = truth, estimate = corrected_pred)

save(brain_age_df_multiwave, file="ABCDMultiwaveBrainAge.Rda")
save(brain_age_corrected_df_multiwave, file="ABCDMultiwaveBrainAgeCorrected.Rda")

```