### 2022-06-07 PaperAnalyses&Figures.Rmd

---
title: "Paper Analyses"
author: "Lucy Whitmore"
date: "5/4/2022"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "ggpubr", 
               "apaTables", "MetBrewer",
               "ggseg", "vip", "tidymodels")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)
```
Load Data
```{r}
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BrainagePDSBaseline.Rda")

#abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineBrainAgeCorrected.Rda")

abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BaselineBrainAgeCorrected.Rda")

load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample1FollowUp.Rda")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample2FollowUp.Rda")

#load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/abcd_baseline.Rda")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineBrainAgeCorrected.Rda")

load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample1Baseline.Rda")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample2Baseline.Rda")

```

# Vlad's Model, Baseline Analyses

#Run Models
```{r}
# brainage ~ youth sum
#baseline_puberty_y_sum_1 <- lm(corrected_gap ~ youth_sum, data = sample_1)
#baseline_puberty_y_sum_2 <- lm(corrected_gap ~ youth_sum, data = sample_2)

sample_1_baseline_months <- sample_1_baseline %>% 
  mutate(interview_age_m = interview_age/12)

sample_2_baseline_months <- sample_2_baseline %>% 
  mutate(interview_age_m = interview_age/12)

# brainage ~ youth mean
baseline_puberty_y_mean_1 <- lm(corrected_gap ~ youth_mean + interview_age, data = sample_1_baseline)
baseline_puberty_y_mean_2 <- lm(corrected_gap ~ youth_mean + interview_age, data = sample_2_baseline)

baseline_puberty_y_mean_1_m <- lm(corrected_gap ~ youth_mean + interview_age_m, data = sample_1_baseline_months)
baseline_puberty_y_mean_2_m <- lm(corrected_gap ~ youth_mean + interview_age_m, data = sample_2_baseline_months)

# brainage ~ parent sum
#baseline_puberty_p_sum_1 <- lm(corrected_gap ~ parent_sum, data = sample_1)
#baseline_puberty_p_sum_2 <- lm(corrected_gap ~ parent_sum, data = sample_2)

# brainage ~ parent mean
baseline_puberty_p_mean_1 <- lm(corrected_gap ~ parent_mean, data = sample_1_baseline)
baseline_puberty_p_mean_2 <- lm(corrected_gap ~ parent_mean, data = sample_2_baseline)

baseline_puberty_p_mean_1_m <- lm(corrected_gap ~ parent_mean + interview_age_m, data = sample_1_baseline_months)
baseline_puberty_p_mean_2_m <- lm(corrected_gap ~ parent_mean + interview_age_m, data = sample_2_baseline_months)

#Cognition
baseline_cognition_1 <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected, data=sample_1_baseline)
baseline_cognition_2 <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected, data=sample_2_baseline)

baseline_cognition_1_m <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected + interview_age_m, data=sample_1_baseline_months)
baseline_cognition_2_m <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected + interview_age_m, data=sample_2_baseline_months)
#Cognition fluid
#baseline_cognition_fluid_1 <- lm(corrected_gap ~ nihtbx_fluidcomp_agecorrected, data=sample_1)
#baseline_cognition_fluid_2 <- lm(corrected_gap ~ nihtbx_fluidcomp_agecorrected, data=sample_2)

#Cognition crystallized
#baseline_cognition_crystal_1 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_1)
#baseline_cognition_crystal_2 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_2)
```

## Model Summaries
```{r}
summary(baseline_puberty_y_mean_1)
summary(baseline_puberty_y_mean_2)

summary(baseline_puberty_p_mean_1)
summary(baseline_puberty_p_mean_2)

summary(baseline_cognition_1)
summary(baseline_cognition_2)
```
Model Summaries of Models w/ age as covariate
```{r}
summary(baseline_puberty_y_mean_1_m)
summary(baseline_puberty_y_mean_2_m)

summary(baseline_puberty_p_mean_1_m)
summary(baseline_puberty_p_mean_2_m)

summary(baseline_cognition_1_m)
summary(baseline_cognition__m2)

```

Figure 4A from vlad’s paper
Make categorical variables for puberty/cognition just for visualization purposes
Geom_density, fill will be different categories

Density plots
X axis brain age gap
y axis density
grouped/split/colored by categorical

3 plots: youth-report puberty, parent-report puberty, cognition
```{r}
#Youth-Report Puberty
#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
categorical_pds_baseline <-  brainage_pds_baseline %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         cognition_cat=cut(nihtbx_totalcomp_agecorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")))

#don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)

ggpar((categorical_pds_baseline %>% 
  filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")

#Parent-Report Puberty
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(parent_pds_cat)) %>% 
ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")

#Cognition
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(cognition_cat) & cognition_cat!="-1-1" ) %>% 
  ggdensity(x = "corrected_gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Age-Corrected Cognition")

#Non-bias corrected version
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")

#Parent-Report Puberty
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(parent_pds_cat)) %>% 
ggdensity(x = "gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")

#Cognition
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(cognition_cat) & cognition_cat!="-1-1" ) %>% 
  ggdensity(x = "gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Age-Corrected Cognition")

```

ggseg ones with weightings for the model
New model only

4 panels:
cortical map and bar graph
subcortical map and bar graph

stacked bar plot with right hemisphere followed by left hemisphere where applicable.
^note: how to decide top 10> top 10 by hemisphere and then add other hemisphere, or top 10 by bilateral calculation?

vip package
```{r}
# Calculate variables of importance
abcd_vi <- vi(abcd_model_5_9) #xgb_final_mod, xgboost_abcd_brain_age_mod
abcd_vi

vip(abcd_model_5_9, num_features = 20)  +
  theme_light()

#Subset so that we don't have to rename all?

# Rename columns to fit with ggseg atlas (ugghhh)

brain_regions(dk)
brain_labels(dk)

brain_labels(aseg)

ggseg(abcd_vi)

#Actually plot
#Cortical
abcd_vi %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = Importance))
#Subcortical
abcd_vi %>%
  ggplot() +
  geom_brain(atlas = aseg, 
             position = position_brain(hemi ~ side),
             aes(fill = Importance))
```

Figure 2
Use subsample, otherwise it’ll look super funky

Scatterplot of scan age/chronological age vs predicted age
2 panels: 1 for validation, 1 for analysis sample
1 or 2 extra panels for analysis sample w/ Vlad's model
```{r}
#subsample data
#sample_n()

ggscatter(brainage_pds_baseline, x = "truth", y = "corrected_pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   xlab= "Scan/Chronological Age", ylab = "Predicted Age"
   )

ggscatter(brainage_pds_baseline, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regression line
   add.params = list(color = "blue", fill = "lightgray")
   )

ggscatter(brainage_pds_baseline, x = "truth", y = "corrected_gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"),
   xlab= "Scan/Chronological Age",
   ylab = "Corrected Gap"
   ) +
  scale_y_continuous(breaks = c(-5, -2.5, 0, 2.5, 5), 
                     labels = c(-5, -2.5, 0, 2.5, 5))

```

Checking Follow-up Brainage

Load Data
```{r}
corrected_followup_brainage <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/FollowUpBrainAgeCorrected.Rda")

#corrected_followup_brainage$src_subject_id <- smri_followup$src_subject_id

```

```{r}
ggpar((corrected_baseline_brainage %>% 
  #filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Baseline")

ggpar((corrected_followup_brainage %>% 
  #filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Follow-Up")

ggscatter(corrected_followup_brainage, x = "truth", y = "corrected_pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(corrected_followup_brainage, x = "truth", y = "corrected_gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(corrected_followup_brainage, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

#Fit Metrics

#baseline uncorrected
corrected_baseline_brainage %>%
  metrics(truth = truth, estimate = .pred)
#baseline corrected
corrected_baseline_brainage %>%
  metrics(truth = truth, estimate = corrected_pred)

#follow-up uncorrected
corrected_followup_brainage %>%
  metrics(truth = truth, estimate = .pred)
#follow-up corrected
corrected_followup_brainage %>%
  metrics(truth = truth, estimate = corrected_pred)

```

Check on talapas model output
```{r}
abcd_model_5_9 <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_brain_age_mod_5_9.rds")

abcd_model_4_22 <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_brain_age_mod.rds")
```

Plots & Decriptives for ABCD Model
```{r}
describe(abcd_brainage_baseline)
#describe(abcd_brainage_baseline %>% filter(truth != 9))
#describe(abcd_brainage_baseline %>% filter(between(truth, 9, 9.2)))

#Fit Metrics
#baseline uncorrected
abcd_brainage_baseline %>%
  metrics(truth = truth, estimate = .pred)
#baseline corrected
abcd_brainage_baseline  %>%
  metrics(truth = truth, estimate = corrected_pred)

ggscatter(abcd_brainage_baseline, x = "truth", y = ".pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(abcd_brainage_baseline, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(abcd_brainage_baseline, x = "truth", y = "corrected_pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   xlab = "Scan/Chronological Age", ylab = "Predicted Age"
   )

ggscatter(abcd_brainage_baseline, x = "truth", y = "corrected_gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   xlab = "Scan/Chronological Age",
   ylab= "Corrected Gap"
   )

hist(abcd_brainage_baseline$truth)

```

### Follow-Up Analyses

```{r}
# brainage ~ youth mean
followup_puberty_y_mean_1 <- lm(corrected_gap ~ youth_mean, data = sample_1_followup)
followup_puberty_y_mean_2 <- lm(corrected_gap ~ youth_mean, data = sample_2_followup)

# brainage ~ parent mean
followup_puberty_p_mean_1 <- lm(corrected_gap ~ parent_mean, data = sample_1_followup)
followup_puberty_p_mean_2 <- lm(corrected_gap ~ parent_mean, data = sample_2_followup)

#Cognition crystallized
followup_cognition_crystal_1 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_1_followup)
followup_cognition_crystal_2 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_2_followup)
```

## Model Summaries
```{r}
summary(followup_puberty_y_mean_1)
summary(followup_puberty_y_mean_2)

summary(followup_puberty_p_mean_1)
summary(followup_puberty_p_mean_2)

summary(followup_cognition_crystal_1)
summary(followup_cognition_crystal_2)
```

#### ABCD Model - Baseline
```{r}
# brainage ~ youth sum
abcd_baseline_puberty_y_sum <- lm(corrected_gap ~ youth_sum, data = abcd_baseline)

# brainage ~ youth mean
abcd_baseline_puberty_y_mean <- lm(corrected_gap ~ youth_mean, data = abcd_baseline)

# brainage ~ parent sum
abcd_baseline_puberty_p_sum <- lm(corrected_gap ~ parent_sum, data = abcd_baseline)

# brainage ~ parent mean
abcd_baseline_puberty_p_mean <- lm(corrected_gap ~ parent_mean, data = abcd_baseline)

#Cognition
abcd_baseline_cognition <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected, data=abcd_baseline)

#Cognition fluid
abcd_baseline_cognition_fluid <- lm(corrected_gap ~ nihtbx_fluidcomp_agecorrected, data=abcd_baseline)

#Cognition crystallized
abcd_baseline_cognition_crystal <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=abcd_baseline)
```

## Model Summaries
```{r}
summary(abcd_baseline_puberty_y_mean)

summary(abcd_baseline_puberty_y_sum)

summary(abcd_baseline_puberty_p_mean)

summary(abcd_baseline_puberty_p_sum)

summary(abcd_baseline_cognition_crystal)

summary(abcd_baseline_cognition_fluid)

summary(abcd_baseline_cognition)
```
ABCD Baseline Model
```{r}
#Youth-Report Puberty
#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
abcd_baseline_categorical_pds <- abcd_baseline %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         cognition_cat=cut(nihtbx_totalcomp_agecorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")))

#don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)

ggpar((abcd_baseline_categorical_pds %>% 
  filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")

#Parent-Report Puberty
ggpar((abcd_baseline_categorical_pds %>% 
  filter(!is.na(parent_pds_cat)) %>% 
ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")

ggpar((abcd_baseline_categorical_pds %>% 
  filter(!is.na(cognition_cat) & cognition_cat!="-1-1" ) %>% 
  ggdensity(x = "gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Age-Corrected Cognition")
```
Examining correlation between vlad's model and abcd model
```{r}

abcd_brainage_baseline$src_subject_id <- sample_2$src_subject_id

model_compare <- abcd_brainage_baseline %>% 
  rename(.pred_abcd = .pred,
         truth_abcd = truth,
        gap_abcd = gap,
        corrected_pred_abcd = corrected_pred,
        corrected_gap_abcd = corrected_gap) %>% 
    left_join(sample_2)

ggscatter(model_compare, x = "corrected_gap", y = "corrected_gap_abcd",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(model_compare, x = "corrected_pred", y = "corrected_pred_abcd",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

cor.test(model_compare$truth, model_compare$corrected_gap_abcd)

cor.test(model_compare$truth, model_compare$corrected_gap)

#Correlation between same participants across models
cor(model_compare$corrected_gap, model_compare$corrected_gap_abcd)

```

###########
Combined Wave Model
###########
```{r}
#Fit Metrics
#baseline uncorrected
abcd_brainage_multiwave %>%
  metrics(truth = truth, estimate = .pred)
#baseline corrected
abcd_brainage_baseline  %>%
  metrics(truth = truth, estimate = corrected_pred)

ggscatter(corrected_multiwave_brainage, x = "truth", y = ".pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(corrected_multiwave_brainage, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(corrected_multiwave_brainage, x = "truth", y = "corrected_pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   xlab = "Scan/Chronological Age", ylab = "Predicted Age"
   )

ggscatter(corrected_multiwave_brainage, x = "truth", y = "corrected_gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   xlab = "Scan/Chronological Age"
   )

```