### 2022-07-11 DifferenceScores.Rmd

---
title: "ChangeScores"
author: "Lucy Whitmore"
date: "7/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "reshape2",
               "nlme", "lme4",
               "data.table", "psych",
               "parallel","lubridate",
               "ggpubr", "broom", 
               "apaTables", "MetBrewer")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)
```

Load Structural Data
```{r}
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/renamed_smri_followup.Rda")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/renamed_smri_baseline.Rda")

smri_abcd_model_baseline <- renamed_smri_baseline[, 1:180] 
smri_abcd_model_followup <- renamed_smri_followup[, 1:180]

```

Load data
```{r}
#Load smri_baseline

#Load smri_followup

smri_difference <- left_join(smri_abcd_model_baseline, smri_abcd_model_followup, by ="src_subject_id")
#could right join to reduce NAs

```

for loop - doesn't work very well
```{r}
difference1 <- c(1, 2, 3, 4, 5)
difference2 <- c(2, 4, 6, 8, 10)
difference3 <- c(1, 3, 5, 7, 9)
difference4 <- c(1, 2, 3, 4, 5)
difference5 <- c(5, 4, 3, 2, 1)
difference6 <- c(1, 2, 3, 5, 8)

difference_df <- data.frame(difference1, difference2, difference3, difference4, difference5, difference6)
difference_df

for (i in 1:(ncol(difference_df)/2)){
  print(i)
  print(i + ncol(difference_df)/2)
  difference_df[ncol(difference_df)+1] <-  difference_df[i + ncol(difference_df)/2] - difference_df[i]
  #print(dif)
}
```

other version
```{r}
#make empty dataframe that only has ids/numbers
#loop through columns, subtract as usual
#rbind (or append, or join, or whatever) the new difference score column onto the new data frame
#end up with dataframe of only ids and difference scores
diff_df <- data.frame(c("id1", "id2", "id3", "id4", "id5"))

for (i in 1:(ncol(difference_df)/2)){
  #print(i)
  #print(i + ncol(difference_df)/2)
  print((difference_df[i + ncol(difference_df)/2] - difference_df[i]))
  cbind(diff_df, (difference_df[i + ncol(difference_df)/2] - difference_df[i]))  
  #print(dif)
}
diff_df
```

Calculate difference scores format
# Stack Exchange - this works!
```{r}

difference1.x <- c(1, 2, 3, 4, 5)
difference2.x <- c(2, 4, 6, 8, 10)
difference3.x <- c(1, 3, 5, 7, 9)
difference1.y <- c(1, 2, 3, 4, 5)
difference2.y <- c(5, 4, 3, 2, 1)
difference3.y <- c(1, 2, 3, 5, 8)

difference_df_xy <- data.frame(difference1.x, difference2.x, difference3.x, difference1.y, difference2.y, difference3.y)
difference_df_xy

# Stack Exchange - this works!
difference_df_xy %>%
 mutate(across(ends_with(".y"), .names = "{col}_diff") - across(ends_with(".x"))) %>%
 rename_with(~ sub("_\\d+", "", .), ends_with("_diff"))

#need to subtract, then divide by age gap
```

Calculate difference scores
```{r}
smri_difference_calculated <- smri_difference %>%
 select(-c(subjectkey.x,subjectkey.y, sex.x,sex.y, eventname.x,eventname.y)) %>% 
 mutate(across(ends_with(".y"), .names = "{col}_diff") - across(ends_with(".x"))) %>%
 rename_with(~ sub("_\\d+", "", .), ends_with("_diff"))

#need to subtract, then divide by age gap
```

Test to make sure this worked
```{r}
smri_difference_calculated$interview_age.y[1] - smri_difference_calculated$interview_age.x[1]
smri_difference_calculated$interview_age.y_diff[1]

smri_difference_calculated$FS_L_Fusiform_Area.y[1] - smri_difference_calculated$FS_L_Fusiform_Area.x[1]
smri_difference_calculated$FS_L_Fusiform_Area.y_diff[1]

#test with NA
smri_difference_calculated$interview_age.y[3] - smri_difference_calculated$interview_age.x[3]
smri_difference_calculated$interview_age.y_diff[3]

smri_difference_calculated$FS_L_Fusiform_Area.y[3] - smri_difference_calculated$FS_L_Fusiform_Area.x[3]
smri_difference_calculated$FS_L_Fusiform_Area.y_diff[3]

#stopifnot(smri_difference_calculated$FS_L_Fusiform_Area.y[3] - smri_difference_calculated$FS_L_Fusiform_Area.x[3] == smri_difference_calculated$FS_L_Fusiform_Area.y_diff[3])

```

Prep for training
```{r}
# Select only difference columns
smri_difference_calculated_subset <- smri_difference_calculated %>% 
  select(src_subject_id, ends_with("diff")) 
  #divide all brain columns by interview age

smri_difference_calculated_scaled<- smri_difference_calculated_subset[,-(1:2)] %>% sapply(`/`, smri_difference_calculated_subset[,2])

smri_difference_calculated_scaled$src_subject_id <- smri_difference_calculated_subset$src_subject_id

smri_difference_calculated_scaled$interview_age_diff <- smri_difference_calculated_subset$interview_age.y_diff

smri_difference_calculated_scaled <- smri_difference_calculated_scaled
select(src_subject_id, interview_age_diff, everything())

# Should we be predicting the change in age?
set.seed(42)
smri_difference_calculated <- smri_difference_calculated %>% 
#split for testing
select(-c(src_subject_id))

```