### 2023-03-01 DiffScoreModel.Rmd
Updated bias metrics, updated names

---
title: "DiffScoreModel"
author: "Lucy Whitmore"
date: "7/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "reshape2",
               "nlme", "lme4",
               "data.table", "psych",
               "parallel","lubridate",
               "ggpubr", "broom", 
               "apaTables", "MetBrewer", "beepr", "doParallel", "tictoc", "rsample", "tidymodels")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)

packageUrl <- "https://cran.r-project.org/src/contrib/Archive/xgboost/xgboost_1.0.0.2.tar.gz"

packageUrl <- "https://cran.r-project.org/src/contrib/Archive/xgboost/xgboost_1.5.2.1.tar.gz"

#packageUrl <- "~/Downloads/xgboost_1.0.0.2.tar.gz"

# You then install this version of the package using
install.packages(packageUrl, repos = NULL, type = 'source')
```

Load data
```{r}
#load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/sample_test_diff.Rda")
#load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/sample_train_diff.Rda")

load("~/Documents/FYP/sample_test_diff.Rda")
load("~/Documents/FYP/sample_train_diff.Rda")
```

Load model
```{r}
#abcd_brainage_diff_model <- readRDS("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/xgboost_abcd_diff_brain_age_mod.rds")

#abcd_brainage_diff_model <- xgboost::xgb.load(readRDS("xgboost_abcd_diff_brain_age_mod"))

abcd_brainage_diff_model <- readRDS("xgboost_abcd_diff_brain_age_mod_updated.rds")

#https://stackoverflow.com/questions/73506085/old-json-xgboost-model-format-will-be-discontinued-in-version-2-3-warning

#abcd_brainage_diff_model_complete <- xgb.Booster.complete(abcd_brainage_diff_model)

#abcd_brainage_multiwave_model <- readRDS("~/Documents/FYP/xgboost_abcd_multiwave_brain_age_mod.rds")

#xgboost::xgb.load(readRDS("xgboost_abcd_diff_brain_age_mod.rds"))

#xgboost::xgb.load(readRDS("xgboost_abcd_multiwave_brain_age_mod.rds"))

```

Predict Brain Age (baseline)
```{r}
sample_test_diff <- sample_test_diff %>% 
  select(-c(src_subject_id, interview_age.y_diff, interview_age.x, interview_age.y))

brain_age_df_diff <- abcd_brainage_diff_model %>%
  predict(new_data = sample_test_diff) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = sample_test_diff$age_midpoint
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

```

Evaluate accuracy
```{r}
# compute common performance metrics: mae, rsq, rmse
brain_age_df_diff  %>%
  metrics(truth = truth, estimate = .pred)

```

Optional bias correction
```{r}
# bias prediction method, steps in comments,
# previously determined values hardcoded.

#get validation set
set.seed(42)
df_split_diff <- initial_split(
  sample_train_diff, 
  prop = 0.80,
  # matching age distributions across train and test set
  strata = "interview_age"
)

df_train_diff <- training(df_split_diff)
df_validation_diff <- testing(df_split_diff)

describe(df_train_diff)
describe(df_validation_diff)

#think I need to now run model on validation set??
abcd_brain_age_correction_diff <-
  abcd_brainage_diff_model %>%
  predict(new_data = df_validation_diff) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = df_validation_diff$age_midpoint
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

#get MAE for hold-out 
abcd_brain_age_correction_diff %>%
  metrics(truth = truth, estimate = .pred)

abcd_bias_mod_diff <- lm(.pred ~ truth, data = abcd_brain_age_correction_diff)

# extract intercept and slope
#abcd_bias_mod_diff$coefficients[["(Intercept)"]]
abcd_bias_intercept_diff <- 9.582399
#abcd_bias_mod_diff$coefficients[["truth"]]
abcd_bias_slope_diff <-  0.1240839

# create bias correct data frame
#Baseline
brain_age_corrected_df_diff <- brain_age_df_diff %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - abcd_bias_intercept_diff) / abcd_bias_slope_diff
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

#mine
brain_age_corrected_df_diff %>%
  metrics(truth = truth, estimate = .pred)

brain_age_corrected_df_diff %>%
  metrics(truth = truth, estimate = corrected_pred)

save(brain_age_df_diff, file="ABCDBrainAgeDiff.Rda")
save(brain_age_corrected_df_diff, file="ABCDBrainAgeDiffCorrected.Rda")

```