### 2023-05-30 Difference Scores /ABCDDiffScoreAnalyses_Updated.Rmd

---
title: "DiffScoreAnalyses"
author: "Lucy Whitmore"
date: "12/6/2022"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(vip)
```

Import Data
```{r}
# make sure to move this file to the server
load("/Volumes/devbrainlab/Lucy/BrainAGE/Diff Scores/Corrected_Diff_BrainAGE.Rda.Rda")
#brain_age_corrected_df_diff

```

Descriptive Plots
```{r}
hist(brain_age_diff_corrected_df$corrected_gap)

ggpar((brain_age_diff_corrected_df %>% 
  #filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Difference Scores")

# Actual vs Corrected Predictions, w/ perfect correlation line

diff_pred_plot <- ggplot(brain_age_diff_corrected_df, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#619B8A")+
 # geom_abline(a=0, b=1, color='#233D4D') +
  labs(x = "Scan/Chronological Age (Midpoint)", y ="Predicted Age (Midpoint)") +
  theme_classic() + 
  theme(text = element_text(size = 14))  
#ggsave(filename="diff_truth_corrected_pred_mod1.png",
  #                        width=7, height=4, units='in', dpi=300)

diff_pred_plot

cor(brain_age_diff_corrected_df$truth, brain_age_diff_corrected_df$corrected_pred)

cor.test(brain_age_diff_corrected_df$truth, brain_age_diff_corrected_df$corrected_pred)

# uncorrected 
ggplot(brain_age_diff_corrected_df, aes(x=truth, y=.pred)) + 
  geom_jitter(color="#619B8A")+
 # geom_abline(a=0, b=1, color='#233D4D') +
  labs(x = "Scan/Chronological Age (Midpoint)", y ="Predicted Age (Midpoint)") +
  theme_classic() + 
  theme(text = element_text(size = 14))

# Age Bias - truth vs corrected gap
diff_gap_plot <- ggplot(brain_age_diff_corrected_df, aes(x=truth, y=corrected_gap)) + 
  geom_jitter(color="#619B8A")+
  geom_smooth(method=lm, color='#233D4D') +
  labs(x = "Scan/Chronological Age (Midpoint)", y ="Corrected Gap") +
  theme_classic() + 
  theme(text = element_text(size = 14))  
#ggsave(filename="followup_truth_corrected_gap_mod1.png",
  #                        width=7, height=4, units='in', dpi=300)

ggscatter(brain_age_diff_corrected_df, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

#Panel Plots
ggarrange(diff_pred_plot, diff_gap_plot, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)

#ggsave(filename = "pred_bias_mod1.png",
#                          width=10, height=10, units='in', dpi=300)

#Fit Metrics

#baseline uncorrected
brain_age_diff_corrected_df %>%
  metrics(truth = truth, estimate = .pred)
#baseline corrected
brain_age_diff_corrected_df %>%
  metrics(truth = truth, estimate = corrected_pred)

```

Variable Importance
```{r}
final_xgb_diff %>%
  fit(data = df_train) %>%
  extract_fit_parsnip() %>%
  vip(geom = "point")

#abcd_diff_vi <- vi(xgb_final_mod)

#write_csv(abcd_diff_vi, "abcd_diff_vi.csv")

#vip(xgb_final_mod, num_features = 20)  +
#  theme_light()

```