### 2023-09-14 Revisions_Analyses&Figures.Rmd
New script for paper revisions

---
title: "Revisions Analyses & Figures"
author: "Lucy Whitmore"
date: "7/26/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install combat
```{r}
library(devtools)
install_github("jfortin1/neuroCombatData")
install_github("jfortin1/neuroCombat_Rpackage")

library(tidymodels) # machine learning metaverse
library(xgboost) # main engine, needs version 1.0.0.2, see above
library(effectsize)
```

Load everything (uncorrected for scanner)
```{r}
# BrainAGE and Behavioral Data - Vlad's Model, Baseline
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BrainagePDSBaseline.Rda")

# BrainAGE and Behavioral Data - Vlad's Model, Follow-Up
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BrainagePDSFollowUp.Rda")

# BrainAGE and Behavioral Data - ABCD Model, Baseline
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineSample.Rda")
```

Load Structural Data
```{r}
#Load list of included subjects
struc_include_baseline<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:11) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include)) %>% 
  filter(imgincl_t1w_include == 1) %>% 
  filter(eventname=="baseline_year_1_arm_1")

struc_include_followup<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:11) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include)) %>% 
  filter(imgincl_t1w_include == 1) %>% 
  filter(eventname=="2_year_follow_up_y_arm_1")

#Load structural data
#Filter for volume & area measurements

##### Need to also select scanner/site #####
smri_baseline <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
  filter(eventname=="baseline_year_1_arm_1") %>% 
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include_baseline$src_subject_id) %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area")) %>% 
  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))

smri_followup <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
  filter(eventname=="2_year_follow_up_y_arm_1") %>% 
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id %in% struc_include_followup$src_subject_id) %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area")) %>% 
  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))

```

# combat/site standardization
```{r}
# load raw data
# filter for quality data
# keep scanner site

#https://github.com/Jfortin1/neuroCombat_Rpackage/tree/2ccab2b42c59163717f012c70a54d20e8757b1c7
# can realign to virtual site or to reference site

data.harmonized <- neuroCombat(dat=dat, batch=batch) # batch is the site/scanner
# maybe need to add age (and gender?) as a covariate, mod =

# maybe run without empirical bayes because number of features is smaller than number of participants?
#data.harmonized <- neuroCombat(dat=dat, batch=batch, eb=FALSE)

# Load data harmonized using LongCombat - longCombat.R
load("Volumes/devbrainlab/Lucy/BrainAGE/fyp/harmonized_mri_baseline.Rda")
load("Volumes/devbrainlab/Lucy/BrainAGE/fyp/harmonized_mri_followup.Rda")

```

Data Cleaning for harmonized data
```{r}
mod_1 <- readRDS(
  file = here::here(
    "DevelopmentalBrainAge/model/xgboost_9to19_brain_age_mod.rds"))

# check for any columns that aren't in model
names.use <- names(model1_features_baseline)[!(names(model1_features_baseline) %in% mod_1$fit$feature_names)]

# baseline
# remove those columns
model1_features_baseline_clean <- model1_features_baseline %>% 
  select(-c(FS_R_Cerebellum_Cort_Vol, interview_age))

# reorder column names to match order of feature names
newdata_baseline = as.matrix(model1_features_baseline_clean)
newdata_baseline = newdata_baseline[ , mod_1$fit$feature_names]

### Follow-Up
# remove those columns
model1_features_followup_clean <- model1_features_followup %>% 
  select(-c(FS_R_Cerebellum_Cort_Vol, interview_age))

# reorder column names to match order of feature names
newdata_followup = as.matrix(model1_features_followup_clean)
newdata_followup = newdata_followup[ , mod_1$fit$feature_names]
```

# rerun predictions on harmonized data
```{r}
# drop scanner site variable from data

# predict brainage - baseline, model 1

# Predict
# This prediction method gives the output as a vector, we'll fix that later on
brainage_mod1_baseline <-
  mod_1$fit %>%
  predict(newdata = newdata_baseline)

# Calculate gap
# Turn the vector into a dataframe, rename the prediction column, add back the column with chronological age, and calculate the gap
brain_age_df_mod1_baseline <- as.data.frame(brainage_mod1_baseline) %>%
  rename(.pred = brainage_mod1_baseline) %>% 
  mutate(
    # provide the chronological age at time of scan
    truth = model1_features_baseline$interview_age) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

# bias correction
# Slope and intercept are hardcoded
bias_intercept <- 6.41 # xgb_bias_mod$coefficients[["(Intercept)"]]
bias_slope <- 0.55 # xgb_bias_mod$coefficients[["truth"]]

# create bias corrected data frame
brain_age_df_mod1_baseline_corrected <- brain_age_df_mod1_baseline %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - bias_intercept) / bias_slope
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

# reattach IDs
brain_age_df_mod1_baseline_corrected$src_subject_id <- harmonized_mri_baseline$src_subject_id

# predict brainage - followup, model 1
brainage_mod1_followup <-
  mod_1$fit %>%
  predict(newdata = newdata_followup)

# Calculate gap
# Turn the vector into a dataframe, rename the prediction column, add back the column with chronological age, and calculate the gap
brain_age_df_mod1_followup <- as.data.frame(brainage_mod1_followup) %>%
  rename(.pred = brainage_mod1_followup) %>% 
  mutate(
    # provide the chronological age at time of scan
    truth = model1_features_followup$interview_age) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

# create bias corrected data frame
brain_age_df_mod1_followup_corrected <- brain_age_df_mod1_followup %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - bias_intercept) / bias_slope
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

# reattach IDs
brain_age_df_mod1_followup_corrected$src_subject_id <- harmonized_mri_followup$src_subject_id

# predict brainage - baseline, model 2
#brainage <-
 # fit_workflow %>%
#  predict(new_data = analysis_sample_baseline) %>%
#  mutate(
    # provide the chronological age at time of scan
#    truth = analysis_sample_baseline$interview_age
 # ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
#  mutate(gap = .pred - truth)

# predict brainage - follow-up, model 2 (3?)

#followup_brainage <-
#  fit_workflow_followup %>%
#  predict(new_data = analysis_sample_baseline) %>%
#  mutate(
    # provide the chronological age at time of scan
#    truth = analysis_sample_baseline$interview_age
 # ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
 # mutate(gap = .pred - truth)
```
# model features and importance
```{r}
# vip plot of feature importance
library(vip)

vip(abcd_model_obj_combat, num_features = 20)

vi(abcd_model_obj_combat)

# ggseg importance plots
# Rename columns to fit with ggseg atlas (ugghhh)

# select only name and atlas name, then join with new vip estimates to get around renaming by hand
abcd_vi_renamed <- rio::import("~/Documents/FYP/abcd_vi_renamed.csv") %>% 
  select(Variable, brain_labels_dk, label)

abcd_vi <- left_join(vi(abcd_model_obj_combat), abcd_vi_renamed)

abcd_vi_renamed_dk <- rio::import("~/Documents/FYP/abcd_vi_renamed_dk.csv") %>% 
  select(Variable, label, aseg_label)

abcd_vi_dk <- left_join(vi(abcd_model_obj_combat), abcd_vi_renamed_dk)

#check labels
brain_labels(dk)
brain_labels(aseg)

#Cortical - Volume
dk_vi_vol <- abcd_vi_dk %>%
  filter(grepl('Vol', Variable)) %>% 
  filter(label != "") %>% 
  select(Importance,label)

dk_vi_vol <- data.frame(dk_vi_vol)

ggplot(dk_vi_vol) +
  geom_brain(atlas = dk, position = position_brain(hemi ~ side), aes(fill = Importance))+
  theme_void() +
  scale_fill_gradient(low = '#FDD872', high = '#E76F51') +
  labs(title = "Cortical Volume")
ggsave(filename="abcd_model_dk_vol_importance.png",
                         width=8, height=4, units='in', dpi=300)

#Cortical - Area
dk_vi_area <- abcd_vi_renamed_dk %>%
  filter(grepl('Area', Variable)) %>% 
  filter(label != "") %>% 
  select(Importance,label)

dk_vi_area <- data.frame(dk_vi_area)

ggplot(dk_vi_area) +
  geom_brain(atlas = dk, position = position_brain(hemi ~ side), aes(fill = Importance))+
  theme_void() +
  scale_fill_gradient(low = '#FDD872', high = '#E76F51') +
  labs(title = "Cortical Area")
ggsave(filename="abcd_model_dk_area_importance.png",
                         width=8, height=4, units='in', dpi=300)

#Subcortical
aseg_vi <- abcd_vi_renamed %>%
  filter(label != "") %>% 
  select(Importance,label)

aseg_vi <- data.frame(aseg_vi)

ggplot(aseg_vi) +
  geom_brain(atlas = aseg, side="coronal", aes(fill = Importance))+
  theme_void() +
  scale_fill_gradient(low = '#FDD872', high = '#E76F51')
ggsave(filename="abcd_model_aseg_importance_coronal.png",
                          width=7, height=4, units='in', dpi=300)

ggplot(aseg_vi) +
  geom_brain(atlas = aseg, side="sagittal", aes(fill = Importance))+
  theme_void() +
  scale_fill_gradient(low = '#FDD872', high = '#E76F51')
ggsave(filename="abcd_model_aseg_importance_sagittal.png",
                          width=7, height=4, units='in', dpi=300)

#troubleshooting names
aseg_labs <- brain_labels(aseg) |> 
  as.character()
aseg_vi$label[!aseg_vi$label %in% aseg_labs]
aseg_labs[!aseg_labs %in% aseg_vi$label]

dk_labs <- brain_labels(dk) |> 
  as.character()
dk_vi_area$label[!dk_vi_area$label %in% dk_labs]
dk_vi_vol$label[!dk_vi_vol$label %in% dk_labs]

dk_labs[!dk_labs %in% dk_vi_area$label]
dk_labs[!dk_labs %in% dk_vi_vol$label]

```

# attach pds and cognition measures to BrainAGE estimates
```{r}
youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(5:38) %>% 
  select(-c(interview_date, pds_remote___1, pds_remote___2, pds_remote___3, pds_remote___4, pds_device)) %>% 
  mutate_at(c(5:28), as.numeric)

#Raw scores, parent
parent_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(5:27) %>% 
  select(-c(interview_date, pds_select_language___1)) %>% 
  mutate_at(c(5:21), as.numeric)

#Cognition Scores (Age-Corrected)
cognition <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(src_subject_id,interview_age, eventname, nihtbx_fluidcomp_agecorrected, nihtbx_cryst_agecorrected, nihtbx_totalcomp_agecorrected, nihtbx_totalcomp_uncorrected) %>% 
  mutate_at(c(2, 4:7), as.numeric)

#two or more missing values = exclude
#PDS mean score is a continuous value calculated by averaging across PDS items following Herting et al. (2021). Items on the PDS were considered missing if participants answered, “I don’t know” or “refuse to answer”, if the response was left blank, or if the response value was outside of the expected range (1 to 4). Participants with two or more missing answers were excluded from the PDS score calculation (following recommendations from (Herting et al., 2021)

### Youth-report, removes everyone w/o complete data
youth_pds[youth_pds == 777] <- NA
youth_pds[youth_pds == 999] <- NA

# Multiwave
youth_pds_multiwave <- youth_pds %>% 
  filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1") %>% 
  mutate(youth_sum = case_when(sex == "M" ~ #male
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_m4_y + pds_m5_y),  
                         sex == "F" ~ #female
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_f4_2_y + pds_f5_y))) %>% 
  mutate(youth_mean = youth_sum/5)  %>% 
  select(src_subject_id, eventname, sex, youth_sum, youth_mean)

youth_pds_baseline <- youth_pds_multiwave %>% 
  filter(eventname == "baseline_year_1_arm_1") %>% 
  select(-eventname)

youth_pds_followup <- youth_pds_multiwave %>% 
  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
  select(-eventname)

#### Parent-report, removes everyone w/o complete data
parent_pds[parent_pds == 777] <- NA
parent_pds[parent_pds == 999] <- NA

#Baseline
parent_pds_multiwave <- parent_pds %>% 
    filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1") %>% 
    mutate(parent_sum = case_when(sex == "M" ~ #male
                           (pds_1_p + pds_2_p + pds_3_p + pds_m4_p + pds_m5_p),  
                         sex == "F" ~ #female
                           (pds_1_p + pds_2_p + pds_3_p + pds_f4_p + pds_f5b_p))) %>% 
  mutate(parent_mean = parent_sum/5) %>% 
  select(src_subject_id, eventname, sex, parent_sum, parent_mean)

parent_pds_baseline <- parent_pds_multiwave %>% 
    filter(eventname == "baseline_year_1_arm_1") %>% 
  select(-eventname)

parent_pds_followup <- parent_pds_multiwave%>% 
    filter(eventname == "2_year_follow_up_y_arm_1")%>% 
  select(-eventname)

#### Cognition

#Baseline
cognition_baseline <- cognition %>% 
  filter(eventname == "baseline_year_1_arm_1") %>% 
  select(-eventname) %>%
  mutate(cog_zscore = scale(nihtbx_totalcomp_uncorrected))

#### Create Dataframes for Analysis #####
# join brainage, pds, cognition - change names
brainage_pds_baseline_combat <- brain_age_df_mod1_baseline_corrected[, c(6,1:5)] %>% 
  left_join(youth_pds_baseline) %>% 
  left_join(parent_pds_baseline) %>% 
  left_join(cognition_baseline)

brainage_pds_followup_combat <- brain_age_df_mod1_followup_corrected[, c(6,1:5)] %>% 
  left_join(youth_pds_followup) %>% 
  left_join(parent_pds_followup)

save(brainage_pds_baseline_combat, file="brainage_pds_baseline_combat.Rda")
save(brainage_pds_followup_combat, file="brainage_pds_followup_combat.Rda")

# abcd model, baseline
abcd_brainage_pds_baseline_combat <- brain_age_corrected_df_combat %>% 
    left_join(youth_pds_baseline) %>% 
    left_join(parent_pds_baseline) %>% 
    left_join(cognition_baseline)

save(abcd_brainage_pds_baseline_combat, file="abcd_brainage_pds_baseline_combat.Rda")

# abcd model, follow-up
abcd_brainage_pds_followup_combat <- brain_age_corrected_df_combat_followup %>% 
    left_join(youth_pds_followup) %>% 
    left_join(parent_pds_followup)

save(abcd_brainage_pds_followup_combat, file="abcd_brainage_pds_followup_combat.Rda")

load("~/Documents/FYP/abcd_brainage_pds_followup_combat.Rda")

```

# rerun analyses without splitting
```{r}
## Baseline - Model 1
# get age in months
# mutate(interview_age_m = interview_age/12)

# brainage ~ youth mean
baseline_puberty_y_mod1 <- lm(corrected_gap ~ youth_mean + truth, data = brainage_pds_baseline_combat)

# brainage ~ parent mean
baseline_puberty_p_mod1 <- lm(corrected_gap ~ parent_mean + truth, data = brainage_pds_baseline_combat)

#standard scores - check that this is actually the one we're trying to use
baseline_cognition_mod1 <- lm(corrected_gap ~ nihtbx_totalcomp_uncorrected + truth, data=brainage_pds_baseline_combat)

## Followup - Model 1
# mutate(interview_age_m = interview_age/12)
# brainage ~ youth mean
followup_puberty_y_mod1 <- lm(corrected_gap ~ youth_mean + truth, data = brainage_pds_followup_combat)

# brainage ~ parent mean
followup_puberty_p_mod1 <- lm(corrected_gap ~ parent_mean + truth, data = brainage_pds_followup_combat)

## Baseline - Model 2
# mutate(interview_age_m = interview_age/12)
# brainage ~ youth mean
baseline_puberty_y_mod2 <- lm(corrected_gap ~ youth_mean + truth, data = abcd_brainage_pds_baseline_combat)

# brainage ~ parent mean
baseline_puberty_p_mod2 <- lm(corrected_gap ~ parent_mean + truth, data = abcd_brainage_pds_baseline_combat)

#standard scores - check that this is actually the one we're trying to use
baseline_cognition_mod2 <- lm(corrected_gap ~ nihtbx_totalcomp_uncorrected + truth, data=abcd_brainage_pds_baseline_combat)

## Followup - Model 2
# mutate(interview_age_m = interview_age/12)
# brainage ~ youth mean
followup_puberty_y_mod2 <- lm(corrected_gap ~ youth_mean + truth, data = abcd_brainage_pds_followup_combat)

# brainage ~ parent mean
followup_puberty_p_mod2 <- lm(corrected_gap ~ parent_mean + truth, data = abcd_brainage_pds_followup_combat)

```

# model summaries
```{r}
summary(baseline_puberty_y_mod1)
summary(baseline_puberty_p_mod1)
summary(baseline_cognition_mod1)

summary(followup_puberty_y_mod1)
summary(followup_puberty_p_mod1)

summary(baseline_puberty_y_mod2)
summary(baseline_puberty_p_mod2)
summary(baseline_cognition_mod2)

summary(followup_puberty_y_mod2)
summary(followup_puberty_p_mod2)

```

# maybe make a table of results
```{r}

```

# standardize effect sizes
```{r}
library(effectsize)
# standardize(model) # updated model, refit with standardized data
# standardize_parameters(models) # table of standardized coefficients from model

# Baseline - Model 1
standardize_parameters(baseline_puberty_y_mod1)
standardize_parameters(baseline_puberty_p_mod1)
standardize_parameters(baseline_cognition_mod1)

# Follow-up - Model 1
standardize_parameters(followup_puberty_y_mod1)
standardize_parameters(followup_puberty_p_mod1)

# Baseline - Model 2
standardize_parameters(baseline_puberty_y_mod2)
standardize_parameters(baseline_puberty_p_mod2)
standardize_parameters(baseline_cognition_mod2)

# Followup - Model 2
standardize_parameters(followup_puberty_y_mod2)
standardize_parameters(followup_puberty_p_mod2)
standardize(followup_puberty_p_mod2)
```

# alternate standardization
```{r}
install.packages("lm.beta")
library(lm.beta)

summary(lm.beta(followup_puberty_y_mod2))

```

# sex differences
```{r}
# model with sex as covariate?

```

# Change plot scaling - Model 1
```{r}
# figure 1 and 4, put actual and chronological on the same scale
# maybe will also need to rerun other plots too, for combat + grid plotting

# Baseline
#baseline_pred_mod1_plot <- 
ggplot(brainage_pds_baseline_combat, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#A1C181", alpha=.2) +
  geom_abline(a=0, b=1, color='#233D4D') +
  xlim(5, 20) + # 5, 20
  ylim(5, 20) +
  #geom_smooth(method=lm, color='#233D4D') +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 14))  
#ggsave(filename="baseline_truth_corrected_pred_mod1.png",
            #              width=7, height=4, units='in', dpi=300)

# Follow-Up
ggplot(brainage_pds_followup_combat, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#619B8A", alpha=.2)+
  geom_abline(a=0, b=1, color='#233D4D') +
  xlim(6, 21) + # 5, 20
  ylim(6, 21) +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 14))

```

# Change plot scaling - Model 2
```{r}
ggplot(abcd_brainage_pds_baseline_combat, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#E76F51", alpha=.25)+
  geom_abline(a=0, b=1, color='#233D4D') +
  xlim(2, 18) + # 5, 20
  ylim(2, 18) +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 12))

ggplot(abcd_brainage_pds_followup_combat, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#BDB5D5", alpha=.25)+
  geom_abline(a=0, b=1, color='#233D4D') +
  xlim(1, 22) + # 5, 20
  ylim(1, 22) +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 12))

```

# Fig 1 with line of identity, jitter fixed
```{r}
# fig 1 has different jitter for ab vs. cd. x axis scaling of ab and cd should be equal
# add line of identity to fig 1 to vizualize bias - already done for AB, but CD has a regression line, do they want a 1:1 line on that too?
```

# Plot of model 1 & 2 prediction overlaid (baseline)
```{r}
# overlay model 1 on new predictions in fig 4

# try combining into one dataframe and then grouping by model
# mutate each dataframe to add column for model
# then rbind
both_mods_baseline <- rbind(brainage_pds_baseline_combat %>% mutate(model = 1), abcd_brainage_pds_baseline_combat %>% mutate(model = 2))

both_mods_baseline <- both_mods_baseline %>% 
  mutate(model = as.factor(model))

ggplot(both_mods_baseline, aes(x=truth, y=corrected_pred, group=model, col=model, fill=model)) +
     # geom_point() +
  geom_jitter(alpha=.2) +
  #geom_abline(a=0, b=1, color='#233D4D') +
  #xlim(8, 15) + # 5, 20
 # ylim(8, 15) +
  #geom_smooth(method=lm, color='#233D4D') +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 14))

# change colors
# decide if it should only show the same participants
```
# Plot of model 1 & 2 prediction overlaid (followup)
```{r}
# overlay model 1 on new predictions in fig 4

# try combining into one dataframe and then grouping by model
# mutate each dataframe to add column for model
# then rbind
both_mods_followup <- rbind(brainage_pds_followup_combat %>% mutate(model = 1), abcd_brainage_pds_followup_combat %>% mutate(model = 2))

both_mods_followup <- both_mods_followup %>% 
  mutate(model = as.factor(model))

ggplot(both_mods_followup, aes(x=truth, y=corrected_pred, group=model, col=model, fill=model)) +
     # geom_point() +
  geom_jitter(alpha=.2) +
  #geom_abline(a=0, b=1, color='#233D4D') +
  #xlim(8, 15) + # 5, 20
 # ylim(8, 15) +
  #geom_smooth(method=lm, color='#233D4D') +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 14))

# change colors
# decide if it should only show the same participants
```

### Rerun old figures
```{r}
# Distrubutions of puberty/cognition by model

```

```{r}
# Prediction Plots

```

```{r}
# Model feature - baseline

# vip

# ggseg

```

### New longitudinal analyses ###

figure out how many baseline training v analysis are in follow-up
```{r}
# check how many from training sample are in follow-up

renamed_harmonized_mri %>% 
  filter(eventname == "2_year_follow_up_y_arm_1")  %>% 
  select(-c(eventname, mri_info_deviceserialnumber)) %>% 
  filter(src_subject_id %in% model2_features$src_subject_id)

#3,739

# check how many from analysis are in follow-up

renamed_harmonized_mri %>% 
  filter(eventname == "2_year_follow_up_y_arm_1")  %>% 
  select(-c(eventname, mri_info_deviceserialnumber)) %>% 
  filter(src_subject_id %in% model2_analysis$src_subject_id)

# 3,693

# total: 7432/7695
```

```{r}
# correlation between gaps from waves, or something more complicated?
# maybe multilevel model so that it's nested by participant? - 
# correlation for vlad's models

#correlation for abcd models

# figure - probably spaghetti plot, show direction of gap change

# for each set of models, look at change in brainage vs chronological age?

```