### 2023-11-16 PaperAnalyses&Figures.Rmd
Cleaned up original script, and added code for descriptive statistics. For revision script, updated code for variable importance, edited figures to save in tiff format and arrange in proper formatting for manuscript. In model training script, updated model names for clarity after incorporation of ComBat pipeline.

---
title: "Paper Analyses"
author: "Lucy Whitmore"
date: "5/4/2022"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "ggpubr", 
               "apaTables", "MetBrewer",
               "ggseg", "vip", "tidymodels", "psych")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)
```
Load Data
```{r}
# BrainAGE and Behavioral Data - Vlad's Model, Baseline
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BrainagePDSBaseline.Rda")

# BrainAGE and Behavioral Data - Vlad's Model, Follow-Up
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BrainagePDSFollowUp.Rda")

# BrainAGE and Behavioral Data - ABCD Model, Baseline
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineSample.Rda")

# ABCD Model Baseline BrainAGE estimates
abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineBrainAgeCorrected.Rda")

#abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BaselineBrainAgeCorrected.Rda")

#load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/abcd_baseline.Rda")
#load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineBrainAgeCorrected.Rda")

load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/FollowUpBrainAgeCorrected.Rda")

load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample1Baseline.Rda")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample2Baseline.Rda")

load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample1FollowUp.Rda")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/Sample2FollowUp.Rda")

# Load Demographics
demographics <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/pdem02.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(src_subject_id, contains("demo_race"), demo_ethn_v2, demo_ethn2_v2, demo_comb_income_v2)

  #join once by baseline IDs, then by follow-up (only recorded at one wave)
demographics_baseline <- left_join(brainage_pds_baseline, demographics)

demographics_followup <- left_join(brainage_pds_followup, demographics)

demographics_baseline_abcdmodel <- left_join(abcd_baseline, demographics)

demographics_long <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_lpds01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(src_subject_id, eventname, demo_comb_income_v2_l)

demographics_baseline <- left_join(demographics_baseline, demographics_long %>% filter(eventname == "baseline_year_1_arm_1"))

demographics_followup <- left_join(demographics_followup, demographics_long %>% filter(eventname == "2_year_follow_up_y_arm_1"))

demographics_baseline_abcdmodel <- left_join(demographics_baseline_abcdmodel, demographics_long %>% filter(eventname == "baseline_year_1_arm_1"))
```

**Vlad's Model, Baseline Analyses**

Run Models
```{r}
# brainage ~ youth sum
#baseline_puberty_y_sum_1 <- lm(corrected_gap ~ youth_sum, data = sample_1)
#baseline_puberty_y_sum_2 <- lm(corrected_gap ~ youth_sum, data = sample_2)

sample_1_baseline_months <- sample_1_baseline %>% 
  mutate(interview_age_m = interview_age/12)

sample_2_baseline_months <- sample_2_baseline %>% 
  mutate(interview_age_m = interview_age/12)

# brainage ~ youth mean
baseline_puberty_y_mean_1 <- lm(corrected_gap ~ youth_mean, data = sample_1_baseline)
baseline_puberty_y_mean_2 <- lm(corrected_gap ~ youth_mean, data = sample_2_baseline)

baseline_puberty_y_mean_1_m <- lm(corrected_gap ~ youth_mean + interview_age_m, data = sample_1_baseline_months)
baseline_puberty_y_mean_2_m <- lm(corrected_gap ~ youth_mean + interview_age_m, data = sample_2_baseline_months)

#baseline_puberty_y_mean_1_pred <- lm(corrected_pred ~ youth_mean, data = sample_1_baseline)
#baseline_puberty_y_mean_2_pred <- lm(corrected_pred ~ youth_mean, data = sample_2_baseline)

#baseline_puberty_y_mean_1_m_pred <- lm(corrected_pred ~ youth_mean + interview_age_m, data = sample_1_baseline_months)
#baseline_puberty_y_mean_2_m_pred <- lm(corrected_pred ~ youth_mean + interview_age_m, data = sample_2_baseline_months)

# brainage ~ parent sum
#baseline_puberty_p_sum_1 <- lm(corrected_gap ~ parent_sum, data = sample_1)
#baseline_puberty_p_sum_2 <- lm(corrected_gap ~ parent_sum, data = sample_2)

# brainage ~ parent mean
baseline_puberty_p_mean_1 <- lm(corrected_gap ~ parent_mean, data = sample_1_baseline)
baseline_puberty_p_mean_2 <- lm(corrected_gap ~ parent_mean, data = sample_2_baseline)

baseline_puberty_p_mean_1_m <- lm(corrected_gap ~ parent_mean + interview_age_m, data = sample_1_baseline_months)
baseline_puberty_p_mean_2_m <- lm(corrected_gap ~ parent_mean + interview_age_m, data = sample_2_baseline_months)

#Cognition
baseline_cognition_1 <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected, data=sample_1_baseline)
baseline_cognition_2 <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected, data=sample_2_baseline)

baseline_cognition_1_m <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected + interview_age_m, data=sample_1_baseline_months)
baseline_cognition_2_m <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected + interview_age_m, data=sample_2_baseline_months)

#standard scores
baseline_cognition_standard_1_m <- lm(corrected_gap ~ nihtbx_totalcomp_uncorrected + interview_age_m, data=sample_1_baseline_months)
baseline_cognition_standard_2_m <- lm(corrected_gap ~ nihtbx_totalcomp_uncorrected + interview_age_m, data=sample_2_baseline_months)

#z-scores
baseline_cognition_zscore_1_m <- lm(corrected_gap ~ cog_zscore + interview_age_m, data=sample_1_baseline_months)
baseline_cognition_zscore_2_m <- lm(corrected_gap ~ cog_zscore + interview_age_m, data=sample_2_baseline_months)

#Cognition fluid
#baseline_cognition_fluid_1 <- lm(corrected_gap ~ nihtbx_fluidcomp_agecorrected, data=sample_1)
#baseline_cognition_fluid_2 <- lm(corrected_gap ~ nihtbx_fluidcomp_agecorrected, data=sample_2)

#Cognition crystallized
#baseline_cognition_crystal_1 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_1)
#baseline_cognition_crystal_2 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_2)
```

Model Summaries
```{r}
summary(baseline_puberty_y_mean_1)
summary(baseline_puberty_y_mean_2)

#summary(baseline_puberty_y_mean_1_pred)
#summary(baseline_puberty_y_mean_2_pred)

summary(baseline_puberty_p_mean_1)
summary(baseline_puberty_p_mean_2)

summary(baseline_cognition_1)
summary(baseline_cognition_2)

summary(baseline_cognition_standard_1)
summary(baseline_cognition_standard_2)

summary(baseline_cognition_zscore_1)
summary(baseline_cognition_zscore_2)

```

Model Summaries of Models w/ age as covariate
```{r}
summary(baseline_puberty_y_mean_1_m)
summary(baseline_puberty_y_mean_2_m)

summary(baseline_puberty_p_mean_1_m)
summary(baseline_puberty_p_mean_2_m)

summary(baseline_cognition_1_m)
summary(baseline_cognition_2_m)

summary(baseline_cognition_standard_1_m)
summary(baseline_cognition_standard_2_m)

```

Descriptive Stats
```{r}
describe(sample_1_baseline_months)

describe(sample_2_baseline_months)

nrow(sample_1_baseline_months %>% filter(sex == "F"))

nrow(sample_2_baseline_months %>% filter(sex == "F"))

```

Basic Prediction Plots
```{r}

df <- ggpredict(baseline_puberty_y_mean_1_m, terms = c("youth_mean","interview_age_m"))
ggplot(df, aes(x, predicted)) + 
    geom_line(aes(linetype=group, color=group)) +
    geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=group), alpha=0.15) +
    scale_linetype_manual(values = c("solid", "dashed", "dotted"))

plot_model(baseline_puberty_y_mean_1_m, type="pred", terms = "youth_mean", title = "Predicted Gap", colors = "#cc5968") +
  xlab("Youth-Report Puberty") +
  ylab("Predicted Gap") +
  theme_sjplot()

plot_model(baseline_puberty_y_mean_2_m, type="pred", terms = "youth_mean", title = "Predicted Gap") +
  xlab("Youth-Report Puberty") +
  ylab("Predicted Gap") +
  theme_sjplot()

plot_model(baseline_puberty_p_mean_1_m, type="pred", terms = "parent_mean", title = "Predicted Gap") +
  xlab("Parent-Report Puberty") +
  ylab("Predicted Gap") +
  theme_sjplot()

plot_model(baseline_puberty_p_mean_2_m, type="pred", terms = "parent_mean", title = "Predicted Gap") +
    xlab("Parent-Report Puberty") +
    ylab("Predicted Gap") +
  theme_sjplot()

plot_model(baseline_cognition_1_m, type="pred", terms = "nihtbx_totalcomp_agecorrected", title = "Predicted Gap")+
   xlab("Age-Corrected Cognition Score") +
    ylab("Predicted Gap") +
  theme_sjplot()

plot_model(baseline_cognition_2_m, type="pred", terms = "nihtbx_totalcomp_agecorrected", title = "Predicted Gap")+
    xlab("Age-Corrected Cognition Score") +
    ylab("Predicted Gap") +
  theme_sjplot()

```

Figure 4A from vlad’s paper
Make categorical variables for puberty/cognition just for visualization purposes
Geom_density, fill will be different categories

Density plots
X axis brain age gap
y axis density
grouped/split/colored by categorical

Color Palettes:
- Darker Green for Model 1 Baseline
#CBDFBD
#BED7AC
#A9CA91
#88B668
- Lighter green for Model 1 follow-uop
#DEE7B1
#D4E09B 
#CAD982
#BDCF63
- Pink for Model 2 baseline
#F5BBA3
#F19C79
#ED865A
#A44A3F

Color Palettes v2
233D4D
E76F51
FCCA46
A1C181
619B8A

Green for M1 Baseline
CCDDBB
A4C185  --
A1C181
7CA352
526D37

Blue for M1 Follow-Up

Red/Yellow for m2 baseline

3 plots: youth-report puberty, parent-report puberty, cognition
```{r}

#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
categorical_pds_baseline <-  brainage_pds_baseline %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         cognition_cat=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")),
         cognition_cat_small=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 92, 108, Inf), labels=c(">-.5", "-.5-.5", ">.5")),
         cog_quart = cut(nihtbx_totalcomp_uncorrected, quantile(nihtbx_totalcomp_uncorrected, na.rm=T), labels=c("0-25", "25-50", "50-75", "75-100")))
#don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)

#Youth-Report Puberty
youth_pds_baseline_mod1_plot <- ggpar((categorical_pds_baseline %>% 
  filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean", palette = c('#3E502A','#7CA352','#A1C181', '#CCDDBB'))),  #,'#526D37'
  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS", font.legend = 8) +
  theme(text = element_text(size = 12)) 
ggsave(filename="baseline_youth_pds_mod1.png",
                          width=7, height=4, units='in', dpi=300)

#Parent-Report Puberty
parent_pds_baseline_mod1_plot <- ggpar((categorical_pds_baseline %>% 
  filter(!is.na(parent_pds_cat)) %>% 
ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean", palette = c('#3E502A','#7CA352','#A1C181', '#CCDDBB'))), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS", font.legend = 8) +
  theme(text = element_text(size = 12)) 
ggsave(filename="baseline_parent_pds_mod1.png",
                          width=7, height=4, units='in', dpi=300)

#Cognition
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(cognition_cat) & cognition_cat!="-1-1" ) %>% 
  ggdensity(x = "corrected_gap", color= "cognition_cat", fill="cognition_cat", add="mean", palette = c('#7CA352',
 '#CCDDBB'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Standardized Cognition")  +
  theme(text = element_text(size = 24))

ggpar((categorical_pds_baseline %>% 
  filter(!is.na(cognition_cat)) %>% 
  ggdensity(x = "corrected_gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Age-Corrected Cognition")

ggpar((categorical_pds_baseline %>% 
  filter(!is.na(cognition_cat_small)) %>% 
  ggdensity(x = "corrected_gap", color= "cognition_cat_small", fill="cognition_cat_small", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Age-Corrected Cognition (.5 sd")

#Cognition Quartiles
cog_baseline_mod1_plot <- ggpar((categorical_pds_baseline %>% 
  filter(!is.na(cog_quart) & cog_quart!="25-50" & cog_quart!="50-75" ) %>% 
  ggdensity(x = "corrected_gap", color= "cog_quart", fill="cog_quart", add="mean", palette = c('#7CA352',
 '#CCDDBB'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Quartiles)", font.legend = 8)  +
  theme(text = element_text(size = 12))

ggsave(filename="baseline_cognition_mod1.png",
                          width=7, height=4, units='in', dpi=300)

ggarrange(youth_pds_baseline_mod1_plot, parent_pds_baseline_mod1_plot, cog_baseline_mod1_plot, 
          labels = c("A", "B", "C"),
          ncol = 2, nrow = 2)

ggsave(filename = "distributions_baseline_mod1.png",
                          width=7, height=5, units='in', dpi=300)

```

Uncorrected Version
```{r}

#Non-bias corrected version
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")

#Parent-Report Puberty
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(parent_pds_cat)) %>% 
ggdensity(x = "gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")

#Cognition
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(cognition_cat) & cognition_cat!="-1-1" ) %>% 
  ggdensity(x = "gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition - Standard Scores")

ggpar((categorical_pds_baseline %>% 
  filter(!is.na(cognition_cat)) %>% 
  ggdensity(x = "gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition - Standard Scores")
```

Demographics for Baseline Sample
```{r}
#white
table(demographics_baseline$demo_race_a_p___10)

#black/african-american
table(demographics_baseline$demo_race_a_p___11)

#native american
table(demographics_baseline$demo_race_a_p___12)

#alaska native
table(demographics_baseline$demo_race_a_p___13)

#native hawaiian
table(demographics_baseline$demo_race_a_p___14)

#guamian
table(demographics_baseline$demo_race_a_p___15)

#samoan
table(demographics_baseline$demo_race_a_p___16)

#other pacific islander
table(demographics_baseline$demo_race_a_p___17)

#asian indian
table(demographics_baseline$demo_race_a_p___18)

#chinese
table(demographics_baseline$demo_race_a_p___19)

#filipino
table(demographics_baseline$demo_race_a_p___20)

#japanese
table(demographics_baseline$demo_race_a_p___21)

#korean
table(demographics_baseline$demo_race_a_p___22)

#vietnamese
table(demographics_baseline$demo_race_a_p___23)

#other asian
table(demographics_baseline$demo_race_a_p___24)

#other race
table(demographics_baseline$demo_race_a_p___25)

#don't know
table(demographics_baseline$demo_race_a_p___99)

#hispanic or latino
table(demographics_baseline$demo_ethn_v2)
```

```{r}
#sex
table(brainage_pds_baseline$sex)

#family income
table(demographics_baseline$demo_comb_income_v2)

#Overall values
#Non-hispanic white
demographics_baseline %>% 
  filter(demo_race_a_p___10 == 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)

#Non-hispanic Black
demographics_baseline %>% 
  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)

#Hispanic
demographics_baseline %>% 
  filter(demo_ethn_v2 == 1)

demographics_baseline %>% 
  filter(((demo_race_a_p___10 == 1 | demo_race_a_p___11 == 1) & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1) & demo_ethn_v2 == 1)

#Asian
demographics_baseline %>% 
  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & ( demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)

#Native American Alaska Native
demographics_baseline %>% 
  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & (demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 )

#Multiracial

# mutate column: true if participant marked any asian category
# mutate column: true if participant marked native american/alaska native
# sum white, black, asian, native american/alaska native, and additional races

demographics_baseline %>% 
  mutate_at(c(18:33), as.numeric) %>% 
  mutate(
    demo_race_asian = case_when((demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) ~ 1,
                            (demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1) ~ 0),
         demo_race_native_american_alaskan = case_when((demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) ~ 1,
                                                       (demo_race_a_p___12 !=1 & demo_race_a_p___13 != 1) ~ 0),
         race_total = rowSums(across(c(demo_race_a_p___10, demo_race_a_p___11, demo_race_a_p___14, demo_race_a_p___15, demo_race_a_p___16, demo_race_a_p___17, demo_race_a_p___25, demo_race_asian, demo_race_native_american_alaskan)), na.rm=T)) %>% 
  filter(race_total > 1)

#additional races
#If the caregiver chose Native Hawaiian, Guamanian, or Samoan, Other Race, or refused to choose a racial category, the youth was placed in the “additional” race group as there were too few youth in each of these categories individually for separate analysis.

demographics_baseline %>% 
  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & (demo_race_a_p___14 == 1 | demo_race_a_p___15 == 1 | demo_race_a_p___16 == 1 | demo_race_a_p___17 == 1 | demo_race_a_p___25 == 1) & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1)

```

Figure 2
Scatterplot of scan age/chronological age vs predicted age
2 panels: 1 for validation, 1 for analysis sample
1 or 2 extra panels for analysis sample w/ Vlad's model
```{r}
#subsample data
#sample_n()

baseline_pred_mod1_plot <- ggplot(brainage_pds_baseline, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#A1C181", alpha=.2) +
  geom_abline(a=0, b=1, color='#233D4D') +
  #geom_smooth(method=lm, color='#233D4D') +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 14))  
ggsave(filename="baseline_truth_corrected_pred_mod1.png",
                          width=7, height=4, units='in', dpi=300)

ggscatter(brainage_pds_baseline, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regression line
   add.params = list(color = "blue", fill = "lightgray")
   )

ggscatter(brainage_pds_baseline, x = "truth", y = "corrected_gap",
   color = "#CBDFBD", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"),
   xlab= "Scan/Chronological Age",
   ylab = "Corrected Gap"
   ) +
  scale_y_continuous(breaks = c(-5, -2.5, 0, 2.5, 5), 
                     labels = c(-5, -2.5, 0, 2.5, 5))

baseline_gap_mod1_plot <- ggplot(brainage_pds_baseline, aes(x=truth, y=corrected_gap)) + 
  geom_jitter(color="#A1C181", alpha=.2)+
  geom_smooth(method=lm, color='#233D4D') +
  labs(x = "Scan/Chronological Age", y ="Corrected Gap") +
  theme_classic() + 
  theme(text = element_text(size = 14))  
ggsave(filename="baseline_truth_corrected_gap_mod1.png",
                          width=7, height=4, units='in', dpi=300)

ggarrange(baseline_pred_mod1_plot, baseline_gap_mod1_plot, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)

ggsave(filename = "pred_bias_baseline_mod1.png",
                          width=10, height=5, units='in', dpi=300)
```

Checking Follow-up Brainage

Load Data
```{r}
corrected_followup_brainage <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/FollowUpBrainAgeCorrected.Rda")

load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/BrainagePDSFollowUp.Rda")

#corrected_followup_brainage$src_subject_id <- smri_followup$src_subject_id

```

Plots
```{r}
ggpar((corrected_baseline_brainage %>% 
  #filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Baseline")

ggpar((corrected_followup_brainage %>% 
  #filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Follow-Up")

# Actual vs Corrected Predictions, w/ perfect correlation line

followup_pred_mod1_plot <- ggplot(brainage_pds_followup, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#619B8A", alpha=.2)+
  geom_abline(a=0, b=1, color='#233D4D') +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 14))  
ggsave(filename="followup_truth_corrected_pred_mod1.png",
                          width=7, height=4, units='in', dpi=300)

# Age Bias - truth vs corrected gap
followup_gap_mod1_plot <- ggplot(brainage_pds_followup, aes(x=truth, y=corrected_gap)) + 
  geom_jitter(color="#619B8A", alpha=.2)+
  geom_smooth(method=lm, color='#233D4D') +
  labs(x = "Scan/Chronological Age", y ="Corrected Gap") +
  theme_classic() + 
  theme(text = element_text(size = 14))  
ggsave(filename="followup_truth_corrected_gap_mod1.png",
                          width=7, height=4, units='in', dpi=300)

ggscatter(corrected_followup_brainage, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

#Panel Plots
ggarrange(baseline_pred_mod1_plot, baseline_gap_mod1_plot, followup_pred_mod1_plot, followup_gap_mod1_plot, 
          labels = c("A", "B", "C", "D"),
          ncol = 2, nrow = 2)

ggsave(filename = "pred_bias_mod1.png",
                          width=10, height=10, units='in', dpi=300)
```

Fit Metrics
```{r}
#baseline uncorrected
corrected_baseline_brainage %>%
  metrics(truth = truth, estimate = .pred)
#baseline corrected
corrected_baseline_brainage %>%
  metrics(truth = truth, estimate = corrected_pred)

#follow-up uncorrected
#corrected_followup_brainage %>%
#  metrics(truth = truth, estimate = .pred)
#follow-up corrected
#corrected_followup_brainage %>%
#  metrics(truth = truth, estimate = corrected_pred)

#follow-up uncorrected
brain_age_corrected_df_followup %>%
  metrics(truth = truth, estimate = .pred)
#follow-up corrected
brain_age_corrected_df_followup%>%
  metrics(truth = truth, estimate = corrected_pred)

# Correlation between gap and age
cor(brainage_pds_baseline$truth, brainage_pds_baseline$corrected_gap)
cor(brainage_pds_followup$truth, brainage_pds_followup$corrected_gap)
```

**Follow-Up Analyses**

```{r}
#Convert ages to months
sample_1_followup_months <- sample_1_followup %>% 
  mutate(interview_age_m = interview_age/12)

sample_2_followup_months <- sample_2_followup %>% 
  mutate(interview_age_m = interview_age/12)

# brainage ~ youth mean
followup_puberty_y_mean_1 <- lm(corrected_gap ~ youth_mean + interview_age_m, data = sample_1_followup_months)
followup_puberty_y_mean_2 <- lm(corrected_gap ~ youth_mean + interview_age_m, data = sample_2_followup_months)

# brainage ~ parent mean
followup_puberty_p_mean_1 <- lm(corrected_gap ~ parent_mean + interview_age_m, data = sample_1_followup_months)
followup_puberty_p_mean_2 <- lm(corrected_gap ~ parent_mean + interview_age_m, data = sample_2_followup_months)

#Cognition crystallized
followup_cognition_crystal_1 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_1_followup)
followup_cognition_crystal_2 <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=sample_2_followup)

# brainage ~ GISH2
```

## Model Summaries
```{r}
# Youth-report puberty
summary(followup_puberty_y_mean_1)
summary(followup_puberty_y_mean_2)

# Parent-report puberty
summary(followup_puberty_p_mean_1)
summary(followup_puberty_p_mean_2)

summary(followup_cognition_crystal_1)
summary(followup_cognition_crystal_2)
```

Descriptive Stats
```{r}
describe(sample_1_followup_months)

describe(sample_2_followup_months)
```

# Density Plots for Follow-Up
```{r}
#Youth-Report Puberty
#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
categorical_pds_followup <-  brainage_pds_followup %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")))

#don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)

youth_pds_followup_mod1_plot <- ggpar((categorical_pds_followup %>% 
  filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean", palette=c('#B4D0C7','#82B0A2', '#619B8A', '#568A7B'))), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS", font.legend=8)  +
  theme(text = element_text(size = 12))

ggsave(filename="followup_youth_pds_mod1.png",
                          width=7, height=4, units='in', dpi=300)

#Parent-Report Puberty
parent_pds_followup_mod1_plot <-ggpar((categorical_pds_followup %>% 
  filter(!is.na(parent_pds_cat)) %>% 
ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean", palette=c('#B4D0C7','#82B0A2', '#619B8A', '#568A7B'))), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS", font.legend=8)  +
  theme(text = element_text(size = 12))

ggsave(filename="followup_parent_pds_mod1.png",
                          width=7, height=4, units='in', dpi=300)

ggarrange(youth_pds_followup_mod1_plot, parent_pds_followup_mod1_plot, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)
ggsave(filename = "distributions_followup_mod1.png",
                          width=10, height=4, units='in', dpi=300)

```

Uncorrected version
```{r}
#Non-bias corrected version
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS")

#Parent-Report Puberty
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(parent_pds_cat)) %>% 
ggdensity(x = "gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean")), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS")

#Cognition
ggpar((categorical_pds_baseline %>% 
  filter(!is.na(cognition_cat) & cognition_cat!="-1-1" ) %>% 
  ggdensity(x = "gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition - Standard Scores")

ggpar((categorical_pds_baseline %>% 
  filter(!is.na(cognition_cat)) %>% 
  ggdensity(x = "gap", color= "cognition_cat", fill="cognition_cat", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition - Standard Scores")

```

Demographics for Follow-Up
```{r}
table(brainage_pds_followup$sex)

#white
table(demographics_followup$demo_race_a_p___10)

#black/african-american
table(demographics_followup$demo_race_a_p___11)

#native american
table(demographics_followup$demo_race_a_p___12)

#alaska native
table(demographics_followup$demo_race_a_p___13)

#native hawaiian
table(demographics_followup$demo_race_a_p___14)

#guamian
table(demographics_followup$demo_race_a_p___15)

#samoan
table(demographics_followup$demo_race_a_p___16)

#other pacific islander
table(demographics_followup$demo_race_a_p___17)

#asian indian
table(demographics_followup$demo_race_a_p___18)

#chinese
table(demographics_followup$demo_race_a_p___19)

#filipino
table(demographics_followup$demo_race_a_p___20)

#japanese
table(demographics_followup$demo_race_a_p___21)

#korean
table(demographics_followup$demo_race_a_p___22)

#vietnamese
table(demographics_followup$demo_race_a_p___23)

#other asian
table(demographics_followup$demo_race_a_p___24)

#other race
table(demographics_followup$demo_race_a_p___25)

#don't know
table(demographics_followup$demo_race_a_p___99)

#hispanic or latino
table(demographics_followup$demo_ethn_v2)

#income
table(demographics_followup$demo_comb_income_v2_l)

#Overall values

#Non-hispanic white
demographics_followup %>% 
  filter(demo_race_a_p___10 == 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)

#Non-hispanic Black
demographics_followup %>% 
  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)

#Hispanic
demographics_followup %>% 
  filter(demo_ethn_v2 == 1)

#Asian
demographics_followup %>% 
  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & ( demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)

#Native American Alaska Native
demographics_followup %>% 
  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & (demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 )

#Multiracial
# mutate column: true if participant marked any asian category
# mutate column: true if participant marked native american/alaska native
# sum white, black, asian, native american/alaska native, and additional races

demographics_followup %>% 
  mutate_at(c(17:32), as.numeric) %>% 
  mutate(
    demo_race_asian = case_when((demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) ~ 1,
                            (demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1) ~ 0),
         demo_race_native_american_alaskan = case_when((demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) ~ 1,
                                                       (demo_race_a_p___12 !=1 & demo_race_a_p___13 != 1) ~ 0),
         race_total = rowSums(across(c(demo_race_a_p___10, demo_race_a_p___11, demo_race_a_p___14, demo_race_a_p___15, demo_race_a_p___16, demo_race_a_p___17, demo_race_a_p___25, demo_race_asian, demo_race_native_american_alaskan)), na.rm=T)) %>% 
  filter(race_total > 1)

#additional races
#If the caregiver chose Native Hawaiian, Guamanian, or Samoan, Other Race, or refused to choose a racial category, the youth was placed in the “additional” race group as there were too few youth in each of these categories individually for separate analysis.

demographics_followup %>% 
  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & (demo_race_a_p___14 == 1 | demo_race_a_p___15 == 1 | demo_race_a_p___16 == 1 | demo_race_a_p___17 == 1 | demo_race_a_p___25 == 1) & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1)

```

**ABCD Model**

Check on talapas model output
```{r}
abcd_model_5_9 <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_brain_age_mod_5_9.rds")

abcd_model_5_9 <- readRDS("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_brain_age_mod_5_9.rds")

abcd_model_multiwave <- readRDS("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_multiwave_brain_age_mod.rds")

xgb.save(abcd_model_5_9, 'abcd_model_5_9.json')

save(abcd_model_5_9, abcd_model_5_9.RData)

xgb.load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_brain_age_mod_5_9.rds")

xgb.serialize(abcd_model_5_9)

xgb.Booster.complete(abcd_model_5_9)

config <- xgb.config(abcd_model_5_9)
print(config)

bst <- xgboost::xgb.Booster("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_brain_age_mod_5_9.rds")

abcd_model_59 <- readRDS("~/Desktop/xgboost_abcd_brain_age_mod_5_9.rds")

abcd_model_4_22 <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/xgboost_abcd_brain_age_mod.rds")
```

Plots & Descriptives for ABCD Model
```{r}
describe(abcd_brainage_baseline)
#describe(abcd_brainage_baseline %>% filter(truth != 9))
#describe(abcd_brainage_baseline %>% filter(between(truth, 9, 9.2)))

#Fit Metrics
#baseline uncorrected
abcd_brainage_baseline %>%
  metrics(truth = truth, estimate = .pred)
#baseline corrected
abcd_brainage_baseline  %>%
  metrics(truth = truth, estimate = corrected_pred)

# Check age v gap
cor(abcd_baseline$truth, abcd_baseline$corrected_gap)

# Plots
ggscatter(abcd_brainage_baseline, x = "truth", y = ".pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(abcd_brainage_baseline, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

# Actual vs Corrected Predictions, w/ perfect correlation line

baseline_pred_mod2_plot <- ggplot(abcd_baseline, aes(x=truth, y=corrected_pred)) + 
  geom_jitter(color="#E76F51", alpha=.25)+
  geom_abline(a=0, b=1, color='#233D4D') +
  labs(x = "Scan/Chronological Age", y ="Predicted Age") +
  theme_classic() + 
  theme(text = element_text(size = 12))  
ggsave(filename="baseline_truth_corrected_pred_mod2.png",
                          width=7, height=4, units='in', dpi=300)

# Age Bias - truth vs corrected gap
baseline_gap_mod2_plot <- ggplot(abcd_baseline, aes(x=truth, y=corrected_gap)) + 
  geom_jitter(color="#E76F51", alpha=.25)+
  geom_smooth(method=lm, color='#233D4D') +
  labs(x = "Scan/Chronological Age", y ="Corrected Gap") +
  theme_classic() + 
  theme(text = element_text(size = 12))  
ggsave(filename="baseline_truth_corrected_gap_mod2.png",
                          width=7, height=4, units='in', dpi=300)

#Panel Plots
ggarrange(baseline_pred_mod2_plot, baseline_gap_mod2_plot,
          labels = c("A", "B"),
          ncol = 2, nrow = 1)

ggsave(filename = "pred_bias_mod2.png",
                          width=10, height=5, units='in', dpi=300)

hist(abcd_brainage_baseline$truth)

```
ggseg ones with weightings for the model
New model only

4 panels:
cortical map and bar graph
subcortical map and bar graph

stacked bar plot with right hemisphere followed by left hemisphere where applicable.
^note: how to decide top 10> top 10 by hemisphere and then add other hemisphere, or top 10 by bilateral calculation?

vip package
```{r}
# Calculate variables of importance
#abcd_vi <- vi(abcd_model_5_9) #xgb_final_mod, xgboost_abcd_brain_age_mod
#abcd_vi

#abcd_vi <- vi(abcd_model_59) #xgb_final_mod, xgboost_abcd_brain_age_mod
#abcd_vi

#xgboost::xgb.importance(model=abcd_brainage_model)
#xgboost::xgb.importance(model=abcd_vi)

abcd_brainage_model <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/xgboost_abcd_brain_age_mod_5_9.rds")

#This works!
#abcd_vi <- vi(abcd_brainage_model) #xgb_final_mod, xgboost_abcd_brain_age_mod
#abcd_vi
#write_csv(abcd_vi, "abcd_vi.csv")

vip(abcd_model_5_9, num_features = 20)  +
  theme_light()

vip(abcd_model_59, num_features = 20)  +
  theme_light()

vip(abcd_brainage_model, num_features = 20)
ggsave(filename="abcd_model_vip.png",
                          width=7, height=4, units='in', dpi=300)

#Subset so that we don't have to rename all?

# Rename columns to fit with ggseg atlas (ugghhh)

abcd_vi_renamed <- rio::import("~/Documents/FYP/abcd_vi_renamed.csv")
abcd_vi_renamed_dk <- rio::import("~/Documents/FYP/abcd_vi_renamed_dk.csv")

#check labels
brain_labels(dk)
brain_labels(aseg)

#Cortical - Volume
dk_vi_vol <- abcd_vi_renamed_dk %>%
  filter(grepl('Vol', Variable)) %>% 
  filter(label != "") %>% 
  select(Importance,label)

dk_vi_vol <- data.frame(dk_vi_vol)

ggplot(dk_vi_vol) +
  geom_brain(atlas = dk, position = position_brain(hemi ~ side), aes(fill = Importance))+
  theme_void() +
  scale_fill_gradient(low = '#FDD872', high = '#E76F51') +
  labs(title = "Cortical Volume")
ggsave(filename="abcd_model_dk_vol_importance.png",
                         width=8, height=4, units='in', dpi=300)

#Cortical - Area
dk_vi_area <- abcd_vi_renamed_dk %>%
  filter(grepl('Area', Variable)) %>% 
  filter(label != "") %>% 
  select(Importance,label)

dk_vi_area <- data.frame(dk_vi_area)

ggplot(dk_vi_area) +
  geom_brain(atlas = dk, position = position_brain(hemi ~ side), aes(fill = Importance))+
  theme_void() +
  scale_fill_gradient(low = '#FDD872', high = '#E76F51') +
  labs(title = "Cortical Area")
ggsave(filename="abcd_model_dk_area_importance.png",
                         width=8, height=4, units='in', dpi=300)

#Subcortical
aseg_vi <- abcd_vi_renamed %>%
  filter(label != "") %>% 
  select(Importance,label)

aseg_vi <- data.frame(aseg_vi)

ggplot(aseg_vi) +
  geom_brain(atlas = aseg, side="coronal", aes(fill = Importance))+
  theme_void() +
  scale_fill_gradient(low = '#FDD872', high = '#E76F51')
ggsave(filename="abcd_model_aseg_importance_coronal.png",
                          width=7, height=4, units='in', dpi=300)

ggplot(aseg_vi) +
  geom_brain(atlas = aseg, side="sagittal", aes(fill = Importance))+
  theme_void() +
  scale_fill_gradient(low = '#FDD872', high = '#E76F51')
ggsave(filename="abcd_model_aseg_importance_sagittal.png",
                          width=7, height=4, units='in', dpi=300)

#troubleshooting names
aseg_labs <- brain_labels(aseg) |> 
  as.character()
aseg_vi$label[!aseg_vi$label %in% aseg_labs]
aseg_labs[!aseg_labs %in% aseg_vi$label]

dk_labs <- brain_labels(dk) |> 
  as.character()
dk_vi_area$label[!dk_vi_area$label %in% dk_labs]
dk_vi_vol$label[!dk_vi_vol$label %in% dk_labs]

dk_labs[!dk_labs %in% dk_vi_area$label]
dk_labs[!dk_labs %in% dk_vi_vol$label]

#'#FCCA46','#F3B5A5', '#E76F51', '#B43718'
```

ABCD Model - Baseline
```{r}
abcd_baseline_months <- abcd_baseline %>% 
  mutate(interview_age_m = interview_age/12)

# brainage ~ youth sum
abcd_baseline_puberty_y_sum <- lm(corrected_gap ~ youth_sum + interview_age_m, data = abcd_baseline_months)

# brainage ~ youth mean
abcd_baseline_puberty_y_mean <- lm(corrected_gap ~ youth_mean+ interview_age_m, data = abcd_baseline_months)
#abcd_baseline_puberty_y_mean_noage <- lm(corrected_gap ~ youth_mean, data = abcd_baseline_months)

# brainage ~ parent sum
abcd_baseline_puberty_p_sum <- lm(corrected_gap ~ parent_sum + interview_age_m, data = abcd_baseline_months)

# brainage ~ parent mean
abcd_baseline_puberty_p_mean <- lm(corrected_gap ~ parent_mean + interview_age_m, data = abcd_baseline_months)
#abcd_baseline_puberty_p_mean_noage <- lm(corrected_gap ~ parent_mean, data = abcd_baseline_months)

#Cognition
abcd_baseline_cognition <- lm(corrected_gap ~ nihtbx_totalcomp_agecorrected+ interview_age_m, data=abcd_baseline_months)

abcd_baseline_cognition_standard <- lm(corrected_gap ~ nihtbx_totalcomp_uncorrected + interview_age_m, data=abcd_baseline_months)

abcd_baseline_cognition_standard_noage <- lm(corrected_gap ~ nihtbx_totalcomp_uncorrected, data=abcd_baseline_months)

#Cognition fluid
#abcd_baseline_cognition_fluid <- lm(corrected_gap ~ nihtbx_fluidcomp_agecorrected, data=abcd_baseline_months)

#Cognition crystallized
#abcd_baseline_cognition_crystal <- lm(corrected_gap ~ nihtbx_cryst_agecorrected, data=abcd_baseline_months)
```

Model Summaries
```{r}
summary(abcd_baseline_puberty_y_mean)

summary(abcd_baseline_puberty_p_mean)

#summary(abcd_baseline_cognition)

summary(abcd_baseline_cognition_standard)
```

Descriptive Stats
```{r}
describe(abcd_baseline_months)
```

Model Plots
```{r}
#Cognition
ggPredict(abcd_baseline_cognition_standard,se=TRUE)

```

ABCD Baseline Model - Density Plots
```{r}
#Youth-Report Puberty
#make categorical variable for puberty
#age-corrected cognition: These are presented as Standard Scores (mean=100, SD=15).
abcd_baseline_categorical_pds <- abcd_baseline %>% 
  mutate(youth_pds_cat=cut(youth_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         parent_pds_cat=cut(parent_mean, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("1","2","3", "4")),
         cognition_cat=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 85, 115, Inf), labels=c(">-1", "-1-1", ">1")),
         cognition_cat_small=cut(nihtbx_totalcomp_uncorrected, breaks=c(-Inf, 92, 108, Inf), labels=c(">-.5", "-.5-.5", ">.5")),
         cog_quart = cut(nihtbx_totalcomp_uncorrected, quantile(nihtbx_totalcomp_uncorrected, na.rm=T), labels=c("0-25", "25-50", "50-75", "75-100")))

#don't z-score, include more than 1 standard deviation and less than 1 standard deviation (ignore within 1 SD)

#Youth-report puberty
youth_pds_baseline_mod2_plot <- ggpar((abcd_baseline_categorical_pds %>% 
  filter(!is.na(youth_pds_cat)) %>% 
  ggdensity(x = "corrected_gap", color= "youth_pds_cat", fill= "youth_pds_cat", add="mean", palette=c('#FCCA46','#F3B5A5', '#E76F51', '#B43718'))), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Youth-Report PDS",  font.legend = 8)  +
  theme(text = element_text(size = 12))

ggsave(filename="baseline_youth_pds_mod2.png",
                          width=7, height=4, units='in', dpi=300)

#Parent-Report Puberty
parent_pds_baseline_mod2_plot <- ggpar((abcd_baseline_categorical_pds %>% 
  filter(!is.na(parent_pds_cat)) %>% 
ggdensity(x = "corrected_gap", color= "parent_pds_cat", fill= "parent_pds_cat", add="mean", palette=c('#FCCA46','#F3B5A5', '#E76F51', '#B43718'))), 
  xlab="Brain Age Gap", ylab="Density", legend.title = "Parent-Report PDS", font.legend = 8)  +
  theme(text = element_text(size = 12))

ggsave(filename="baseline_parent_pds_mod2.png",
                          width=7, height=4, units='in', dpi=300)

ggpar((abcd_baseline_categorical_pds %>% 
  filter(!is.na(cognition_cat)) %>% 
  ggdensity(x = "corrected_gap", color= "cognition_cat", fill="cognition_cat", add="mean", palette=c('#FCCA46',
 '#E76F51'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Standard Scores") +
  theme(text = element_text(size = 24))

ggpar((abcd_baseline_categorical_pds %>% 
  filter(!is.na(cognition_cat)) %>% 
  ggdensity(x = "corrected_gap", color= "cognition_cat_small", fill="cognition_cat_small", add="mean")), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Standard Scores) .5 SD") +
  theme(text = element_text(size = 24))

#Cognition Quartiles
cog_baseline_mod2_plot <- ggpar((abcd_baseline_categorical_pds %>% 
  filter(!is.na(cog_quart) & cog_quart!="25-50" & cog_quart!="50-75" ) %>% 
  ggdensity(x = "corrected_gap", color= "cog_quart", fill="cog_quart", add="mean", palette = c('#FCCA46',
 '#E76F51'))), xlab="Brain Age Gap", ylab="Density", legend.title = "Cognition (Quartiles)", font.legend = 8)  +
  theme(text = element_text(size = 12))

ggsave(filename="baseline_cognition_mod2.png",
                          width=7, height=4, units='in', dpi=300)

ggarrange(youth_pds_baseline_mod2_plot, parent_pds_baseline_mod2_plot, cog_baseline_mod2_plot, 
          labels = c("A", "B", "C"),
          ncol = 2, nrow = 2)
ggsave(filename = "distributions_baseline_mod2.png",
                          width=7, height=5, units='in', dpi=300)
```

Demographics for ABCD Baseline
```{r}
table(demographics_baseline_abcdmodel$sex)

#white
table(demographics_baseline_abcdmodel$demo_race_a_p___10)

#black/african-american
table(demographics_baseline_abcdmodel$demo_race_a_p___11)

#native american
table(demographics_baseline_abcdmodel$demo_race_a_p___12)

#alaska native
table(demographics_baseline_abcdmodel$demo_race_a_p___13)

#native hawaiian
table(demographics_baseline_abcdmodel$demo_race_a_p___14)

#guamian
table(demographics_baseline_abcdmodel$demo_race_a_p___15)

#samoan
table(demographics_baseline_abcdmodel$demo_race_a_p___16)

#other pacific islander
table(demographics_baseline_abcdmodel$demo_race_a_p___17)

#asian indian
table(demographics_baseline_abcdmodel$demo_race_a_p___18)

#chinese
table(demographics_baseline_abcdmodel$demo_race_a_p___19)

#filipino
table(demographics_baseline_abcdmodel$demo_race_a_p___20)

#japanese
table(demographics_baseline_abcdmodel$demo_race_a_p___21)

#korean
table(demographics_baseline_abcdmodel$demo_race_a_p___22)

#vietnamese
table(demographics_baseline_abcdmodel$demo_race_a_p___23)

#other asian
table(demographics_baseline_abcdmodel$demo_race_a_p___24)

#other race
table(demographics_baseline_abcdmodel$demo_race_a_p___25)

#don't know
table(demographics_baseline_abcdmodel$demo_race_a_p___99)

#hispanic or latino
table(demographics_baseline_abcdmodel$demo_ethn_v2)

#income
table(demographics_baseline_abcdmodel$demo_comb_income_v2)

# calculate overall values
#Overall values
#Non-hispanic white
demographics_baseline_abcdmodel %>% 
  filter(demo_race_a_p___10 == 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)

#Non-hispanic Black
demographics_baseline_abcdmodel%>% 
  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)

#Hispanic
demographics_baseline_abcdmodel %>% 
  filter(demo_ethn_v2 == 1)

#Asian
demographics_baseline_abcdmodel %>% 
  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & ( demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) & demo_race_a_p___25 != 1 & demo_ethn_v2 != 1)

#Native American Alaska Native
demographics_baseline_abcdmodel %>% 
  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 != 1 & (demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) & demo_race_a_p___14 != 1 & demo_race_a_p___15 != 1 & demo_race_a_p___16 != 1 & demo_race_a_p___17 != 1 & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1 & demo_race_a_p___25 != 1 )

#Multiracial
#Multiracial

# mutate column: true if participant marked any asian category
# mutate column: true if participant marked native american/alaska native
# sum white, black, asian, native american/alaska native, and additional races

demographics_baseline_abcdmodel %>% 
  mutate_at(c(18:33), as.numeric) %>% 
  mutate(
    demo_race_asian = case_when((demo_race_a_p___18 == 1 | demo_race_a_p___19 == 1 | demo_race_a_p___20 == 1 | demo_race_a_p___21 == 1 | demo_race_a_p___22 == 1 | demo_race_a_p___23 == 1 | demo_race_a_p___24 == 1) ~ 1,
                            (demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1) ~ 0),
         demo_race_native_american_alaskan = case_when((demo_race_a_p___12 ==1 | demo_race_a_p___13 == 1) ~ 1,
                                                       (demo_race_a_p___12 !=1 & demo_race_a_p___13 != 1) ~ 0),
         race_total = rowSums(across(c(demo_race_a_p___10, demo_race_a_p___11, demo_race_a_p___14, demo_race_a_p___15, demo_race_a_p___16, demo_race_a_p___17, demo_race_a_p___25, demo_race_asian, demo_race_native_american_alaskan)), na.rm=T)) %>% 
  filter(race_total > 1)

#additional races
#If the caregiver chose Native Hawaiian, Guamanian, or Samoan, Other Race, or refused to choose a racial category, the youth was placed in the “additional” race group as there were too few youth in each of these categories individually for separate analysis.

demographics_baseline_abcdmodel %>% 
  filter(demo_race_a_p___10 != 1 & demo_race_a_p___11 == 1 & demo_race_a_p___12 != 1 & demo_race_a_p___13 != 1 & (demo_race_a_p___14 == 1 | demo_race_a_p___15 == 1 | demo_race_a_p___16 == 1 | demo_race_a_p___17 == 1 | demo_race_a_p___25 == 1) & demo_race_a_p___18 != 1 & demo_race_a_p___19 != 1 & demo_race_a_p___20 != 1 & demo_race_a_p___21 != 1 & demo_race_a_p___22 != 1 & demo_race_a_p___23 != 1 & demo_race_a_p___24 != 1)

```

Examining correlation between vlad's model and abcd model
```{r}

abcd_brainage_baseline$src_subject_id <- sample_2$src_subject_id

model_compare <- abcd_brainage_baseline %>% 
  rename(.pred_abcd = .pred,
         truth_abcd = truth,
        gap_abcd = gap,
        corrected_pred_abcd = corrected_pred,
        corrected_gap_abcd = corrected_gap) %>% 
    left_join(sample_2)

ggscatter(model_compare, x = "corrected_gap", y = "corrected_gap_abcd",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(model_compare, x = "corrected_pred", y = "corrected_pred_abcd",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

cor.test(model_compare$truth, model_compare$corrected_gap_abcd)

cor.test(model_compare$truth, model_compare$corrected_gap)

#Correlation between same participants across models
cor(model_compare$corrected_gap, model_compare$corrected_gap_abcd)

```

Checking internal consistency of puberty means
```{r}
#Puberty
#Raw scores, youth
youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(5:38) %>% 
  select(-c(interview_date, pds_remote___1, pds_remote___2, pds_remote___3, pds_remote___4, pds_device)) %>% 
  mutate_at(c(5:28), as.numeric)

#Raw scores, parent
parent_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(5:27) %>% 
  select(-c(interview_date, pds_select_language___1)) %>% 
  mutate_at(c(5:21), as.numeric)

#two or more missing values = exclude
#PDS mean score is a continuous value calculated by averaging across PDS items following Herting et al. (2021). Items on the PDS were considered missing if participants answered, “I don’t know” or “refuse to answer”, if the response was left blank, or if the response value was outside of the expected range (1 to 4). Participants with two or more missing answers were excluded from the PDS score calculation (following recommendations from (Herting et al., 2021)

### Youth-report, removes everyone w/o complete data
youth_pds[youth_pds == 777] <- NA
youth_pds[youth_pds == 999] <- NA

#### Parent-report, removes everyone w/o complete data
parent_pds[parent_pds == 777] <- NA
parent_pds[parent_pds == 999] <- NA

# Calculting internal consistency
# Use alpha output

# Baseline, youth-report, male
youth_pds %>% 
  #filter for baseline wave, male, has BrainAGE estimate
  filter(eventname == "baseline_year_1_arm_1", sex == "M", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
  #select values being used for mean calculation (male)
  select(c("pds_ht2_y", "pds_bdyhair_y", "pds_skin2_y","pds_m4_y","pds_m5_y")) %>% 
  psych::omega(nfactors=3)

# Baseline, youth-report, female
youth_pds %>% 
  #filter for baseline wave, female, has BrainAGE estimate
  filter(eventname == "baseline_year_1_arm_1", sex == "F", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
  #select values being used for mean calculation (male)
  select(c("pds_ht2_y", "pds_bdyhair_y", "pds_skin2_y","pds_f4_2_y","pds_f5_y")) %>% 
  psych::omega(nfactors=3)

# Baseline, parent-report, male                           
parent_pds %>% 
  #filter for baseline wave, male, has BrainAGE estimate
  filter(eventname == "baseline_year_1_arm_1", sex == "M", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
  #select values being used for mean calculation (male)
  select(c("pds_1_p", "pds_2_p", "pds_3_p", "pds_m4_p", "pds_m5_p")) %>% 
  psych::omega(nfactors=3)

# Baseline, parent-report, female                           
parent_pds %>% 
  #filter for baseline wave, female, has BrainAGE estimate
  filter(eventname == "baseline_year_1_arm_1", sex == "F", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
  #select values being used for mean calculation (male)
  select(c("pds_1_p", "pds_2_p", "pds_3_p", "pds_f4_p", "pds_f5b_p")) %>% 
  psych::omega(nfactors=3)

# Follow-up, youth-report, male
youth_pds %>% 
  #filter for baseline wave, male, has BrainAGE estimate
  filter(eventname == "2_year_follow_up_y_arm_1", sex == "M", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
  #select values being used for mean calculation (male)
  select(c("pds_ht2_y", "pds_bdyhair_y", "pds_skin2_y","pds_m4_y","pds_m5_y")) %>% 
  psych::omega(nfactors=3)

# Follow-up, youth-report, female
youth_pds %>% 
  #filter for baseline wave, female, has BrainAGE estimate
  filter(eventname == "2_year_follow_up_y_arm_1", sex == "F", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
  #select values being used for mean calculation (male)
  select(c("pds_ht2_y", "pds_bdyhair_y", "pds_skin2_y","pds_f4_2_y","pds_f5_y")) %>% 
  psych::omega(nfactors=3)

# Follow-up, parent-report, male                           
parent_pds %>% 
  #filter for baseline wave, male, has BrainAGE estimate
  filter(eventname == "2_year_follow_up_y_arm_1", sex == "M", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
  #select values being used for mean calculation (male)
  select(c("pds_1_p", "pds_2_p", "pds_3_p", "pds_m4_p", "pds_m5_p")) %>% 
  psych::omega(nfactors=3)

# Follow-up, parent-report, female                           
parent_pds %>% 
  #filter for baseline wave, female, has BrainAGE estimate
  filter(eventname == "2_year_follow_up_y_arm_1", sex == "F", src_subject_id %in% brainage_pds_baseline$src_subject_id) %>% 
  #select values being used for mean calculation (male)
  select(c("pds_1_p", "pds_2_p", "pds_3_p", "pds_f4_p", "pds_f5b_p")) %>% 
  psych::omega(nfactors=3)

```

Check correlation between youth and parent report puberty
```{r}
# baseline youth & parent report
cor(brainage_pds_baseline$youth_mean, brainage_pds_baseline$parent_mean, use="complete.obs")

# follow-up youth and parent report
cor(brainage_pds_followup$youth_mean, brainage_pds_followup$parent_mean, use="complete.obs")

# abcd model baseline
cor(abcd_baseline$youth_mean, abcd_baseline$parent_mean, use="complete.obs")

```

**Combined Wave Model**

```{r}
#Fit Metrics
#baseline uncorrected
abcd_brainage_multiwave %>%
  metrics(truth = truth, estimate = .pred)
#baseline corrected
abcd_brainage_baseline  %>%
  metrics(truth = truth, estimate = corrected_pred)

ggscatter(corrected_multiwave_brainage, x = "truth", y = ".pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(corrected_multiwave_brainage, x = "truth", y = "gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )

ggscatter(corrected_multiwave_brainage, x = "truth", y = "corrected_pred",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   xlab = "Scan/Chronological Age", ylab = "Predicted Age"
   )

ggscatter(corrected_multiwave_brainage, x = "truth", y = "corrected_gap",
   color = "#00AFBB", # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   xlab = "Scan/Chronological Age"
   )

```