### 2023-11-16 UpdatedModels/ABCDModelTraining_FollowUp.Rmd
Added model training script for follow-up data.

---
title: "Train BrainAGE with tidymodels"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, fig.height = 5)
```

```{r}
library(tidymodels)
library(xgboost)
```

```{r}
#load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/training_sample_baseline.Rds")
load("~/Documents/FYP/Revisions/model2_features_followup.Rda")

model2_features_followup_noid <- model2_features_followup %>% 
  select(-c(src_subject_id))

#training_subset_test <- training_sample_baseline[1:100,]

# Model prep: split, preprocessing, CV ------------------------------------

# Train / test split ------------------------------------------------------

set.seed(42)
df_split_followup <- initial_split(
  model2_features_followup_noid, 
  prop = 0.80,
  # matching age distributions across train and test set
  strata = "interview_age"
)
df_train_followup <- training(df_split_followup)
df_validation_followup <- testing(df_split_followup)

```

# Pre-processing setup ----------------------------------------------------

```{r}
# define (what we want to do)
preprocess_recipe_followup <- df_train_followup %>%
  # predict scan age by all brain features
  recipe(interview_age ~ .) %>%
  # remove near zero variance predictors
  step_nzv(all_predictors()) %>%
  prep() # where it all gets calculated

preprocess_recipe_followup

# Apply pre-processing ----------------------------------------------------

# juice() will work with training data, `bake()` to apply this to our test data

# apply on train (gives processed value)
df_train_followup_prep <- juice(preprocess_recipe_followup)

# apply on validation
df_validation_followup_prep <- preprocess_recipe %>% bake(df_validation_followup)

```

```{r}
boost_mod_followup <- boost_tree(
  mode = "regression", 
  trees = 150, 
  tree_depth = tune(), min_n = tune(), loss_reduction = tune(),
  # randomness
  sample_size = tune(), mtry = tune(), 
  # step size
  learn_rate = tune()
) %>%
  set_engine("xgboost", 
             objective = "reg:squarederror")

```

```{r}
set.seed(42)

xgboost_grid_followup <- grid_latin_hypercube(
  min_n(), 
  tree_depth(), 
  loss_reduction(),
  sample_size = sample_prop(),
  # has unknown, finalize with data to find max
  finalize(mtry(), df_train_followup_prep),
  learn_rate(),
  size = 500 
)

```

```{r}
xgb_wf_followup <- workflow() %>%
  add_formula(interview_age ~ .) %>%
  add_model(boost_mod_followup)

xgb_wf_followup
```

```{r}
set.seed(42)

train_cv_followup <- df_train_followup_prep %>%
  vfold_cv(
    v = 10, 
    repeats = 10, 
    strata = interview_age
  )
```

```{r}
doParallel::registerDoParallel()

set.seed(42)

xgb_tuned_results_followup <- tune_grid(
  xgb_wf_followup,
  resamples = train_cv_followup,
  grid = xgboost_grid_followup,
  metrics = metric_set(mae, rmse, rsq),
  control = control_grid(verbose = TRUE,
                         save_pred = TRUE)
)

xgb_tuned_results_followup
```

```{r}
xgb_tuned_results_followup %>%
  # want to minimize MAE
  show_best("mae")

# select parsimonious params within one SE of best model
best_xgb_params_followup <- xgb_tuned_results_followup %>%
  select_by_one_std_err(metric = "mae", maximize = FALSE, tree_depth)

```

Finalize workflow and fit final model
```{r}
final_xgb_combat_followup <- finalize_workflow(
  xgb_wf_followup,
  best_xgb_params_followup
)

final_xgb_combat_followup

fit_workflow_combat_followup <- fit(final_xgb_combat_followup, df_train_followup_prep)

#xgb_final_mod_test <- boost_mod_test %>%
 # finalize_model(best_xgb_params_test) %>%
 # fit(interview_age ~ .,
#      data = df_train_prep)

```

Variable importance, ignore
```{r}
library(vip)

final_xgb_combat_followup %>%
  fit(data = df_train_followup) %>%
  extract_fit_parsnip() %>%
  vip(geom = "point")

```

```{r}
final_res_followup <- last_fit(final_xgb_combat_followup, df_split_followup)

collect_metrics(final_res_followup)
```

Save models
```{r}
# three different save formats just in case
# 1. Rds
save(fit_workflow_combat_followup, file = "fit_workflow_combat_followup.rds")
# 2. rda
save(fit_workflow_combat_followup, file = "fit_workflow_2_combat_followup.rda")
# 3. raw xbg object, which can then be loaded with xgb.load
abcd_model_obj_combat_followup <- fit_workflow_combat_followup$fit$fit$fit
xgb.save(abcd_model_obj_combat_followup, "abcd_model_obj_combat_followup")
```

Model bundling, ignore for now
```{r}
# Testing out bundling
#mod_bundle <- bundle::bundle(fit_workflow)
#save(mod_bundle, file = "xgb_workflow_bundles.rds")

#save(mod_bundle, file = "xgb_workflow.rds")

# in a new R session:
#mod_bundle <- readRDS("path/to/file.rds")
#mod_new <- unbundle(mod_bundle)

#xgb.save.raw(fit_workflow$fit$fit, raw_format = "deprecated")

#model_obj <- fit_workflow$fit$fit$fit
#xgb.save(model_obj, "model_obj")

#xgb.save.raw(model_obj, "model_obj_raw")
```

Predict
```{r}
load("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/Revisions/model2_analysis_followup.Rds")

# combat version
brain_age_df_combat_followup <-
  abcd_model_obj_combat_followup %>%
  predict(newdata = as.matrix(model2_analysis_followup %>% select(-c(src_subject_id, interview_age))))

brain_age_df_combat_followup <- as.data.frame(brain_age_df_combat_followup) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = model2_analysis_followup$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

#workflow_predict <- predict(fit_workflow, analysis_sample)

brain_age_df_workflow_combat_followup <- fit_workflow_combat_followup %>%
  predict(new_data = model2_analysis_followup) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = model2_analysis_followup$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)
```

Bias Correction
```{r}
#run model on validation set
abcd_brain_age_correction_followup <-
  fit_workflow_combat_followup %>%
  predict(new_data = df_validation_followup) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = df_validation_followup$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

#get MAE for hold-out 
abcd_brain_age_correction_followup %>%
  metrics(truth = truth, estimate = .pred)

abcd_bias_mod_followup <- lm(.pred ~ truth, data = abcd_brain_age_correction_followup)

# extract intercept and slope
#
abcd_bias_intercept_followup <- abcd_bias_mod_followup$coefficients[["(Intercept)"]]

abcd_bias_slope_followup <-  abcd_bias_mod_followup$coefficients[["truth"]]

# create bias correct data frame
#Baseline
brain_age_corrected_df_combat_followup <- brain_age_df_workflow_combat_followup %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - abcd_bias_intercept_followup) / abcd_bias_slope_followup
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

```

Clean up & save
```{r}
# add IDs back
brain_age_corrected_df_combat_followup$src_subject_id <- model2_analysis_followup$src_subject_id

brain_age_corrected_df_combat_followup <- brain_age_corrected_df_combat_followup %>% 
  select(6, 1:5)

# save dataframe
save(brain_age_corrected_df_combat_followup, file = "brain_age_corrected_df_combat_followup.Rda")

```