### 2023-11-16 FreeSurferSetupBrainAGE.Rmd
Added script for converting names from freesurfer output, updated readme.

---
title: "NavNetworks BrainAGE"
author: "Lucy Whitmore"
date: "4/3/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <-  c("tidyverse",
               "nlme", "lme4",
               "ggpubr", "lubridate", "tidymodels")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
lapply(packages, library, character.only = TRUE)
```

Import & combine data
```{r}
left_area <- rio::import("left_area.txt")  %>% 
  rename(subid = lh.aparc.area)

right_area <- rio::import("right_area.txt")   %>% 
  rename(subid = rh.aparc.area)

right_volume <- rio::import("right_volume.txt")   %>% 
  rename(subid = rh.aparc.volume)

left_volume<- rio::import("left_volume.txt")   %>% 
  rename(subid = lh.aparc.volume)

aseg_volume <- rio::import("aseg_volume.txt")  %>% 
  rename(subid = "Measure:volume")

```

rename columns
```{r}
#aseg_volume just needs _Vol added
#regex mutate column names
#add a column for aseg vs aparc and grep based on that

# replace cortex with cort
# ventricle with vent

left_volume <- left_volume %>% rename_with(~str_replace(., "_volume", "_GrayVol"))

right_volume <- right_volume %>% rename_with(~str_replace(., "_volume", "_GrayVol"))

names(aseg_volume)[-1] <- names(aseg_volume)[-1] %>% 
  paste0("_Vol")

names(aseg_volume)[-1] <- paste0("FS_", names(aseg_volume)[-1])
```

```{r}
brain_area <- left_join(left_area, right_area, by = "subid")

brain_volume <- left_join(left_volume, right_volume, by = "subid")

brain_aparc <- left_join(brain_area, brain_volume, by = "subid")

brain <- left_join(brain_aparc, aseg_volume, by ="subid")
```

```{r}

brain_names <- brain

brain_names <- brain_names %>% 
  rename_with(~str_replace_all(., c("lh_" = "FS_L_", "rh_" = "FS_R_", "FS_Left-" = "FS_L_", "FS_Right-" = "FS_R_", "FS_lh" = "FS_L_", "FS_rh" = "FS_R_", "Cortex" = "Cort",
                    "Ventricle" = "Vent", "-" = "_")))

colnames(brain_names) <- gsub("(?<=^|_)([a-z])", "\\U\\1", colnames(brain_names), perl=T)

```

add extra columns to brain, pivot and cbind to pivoted training set
```{r}
#brain_copy <- brain
training_copy <- training_sample_baseline
namevector <- names(brain[177:211])
training_copy[,namevector] <- NA

row_num <- seq(1,211)
#feature_df <- as.data.frame(row_num, names(brain_copy), names(training_sample_baseline))

feature_df <- as.data.frame(cbind(row_num, names(brain), names(training_copy))) %>% 
  select(V2, V3) %>% 
  rename(nav_names = V2,
         model_names = V3)

#make values from 140 to end in nav NAs
feature_df$model_names[177:211] <- NA
```

regex test
- clean1 and 2 used because subsequent gsub statements will override the others, nesting didn't work. 
```{r}

#change lh_ to FS_L_, rh_ to FS_R_
#group$group.no.e <- gsub("e", "", group$group)
feature_df$nav_names_clean <- gsub("lh_", "FS_L_", feature_df$nav_names)
feature_df$nav_names_clean <- gsub("rh_", "FS_R_", feature_df$nav_names_clean)

#replace Left- with FS_L_
feature_df$nav_names_clean <- gsub("FS_Left-", "FS_L_", feature_df$nav_names_clean)
feature_df$nav_names_clean <- gsub("FS_Right-", "FS_R_", feature_df$nav_names_clean)

# replace cortex with cort
feature_df$nav_names_clean <- gsub("Cortex", "Cort", feature_df$nav_names_clean)
# ventricle with vent
feature_df$nav_names_clean <- gsub("Ventricle", "Vent", feature_df$nav_names_clean)
# - with _
feature_df$nav_names_clean <- gsub("-", "_", feature_df$nav_names_clean)

#capitalize first letter after _
feature_df$nav_names_clean <- gsub("(?<=^|_)([a-z])", "\\U\\1", feature_df$nav_names_clean, perl=TRUE)

```

find missing variables -old
```{r}
# Note: this is just one way to do this, and compares the new data to the training sample, which is not hosted publicaly. Comparing to feature_list.txt might be a workaround.

# get columns in analysis sample that aren't in brain
hcp_diff <- as.data.frame(setdiff(names(analysis_sample_baseline), names(brain)))

# get columns in analysis sample that aren't in brain
nav_diff <- as.data.frame(setdiff(names(brain), names(analysis_sample_baseline)))

max_length <- max(length(hcp_diff), length(nav_diff))

#set length of each vector equal to max length
length(hcp_diff) <- max_length                      
length(nav_diff) <- max_length

#cbind the two vectors together
cbind(hcp_diff, nav_diff)

diff_df <- as.data.frame(c(setdiff(names(analysis_sample_baseline), names(brain)), setdiff(names(brain), names(analysis_sample_baseline))))
```

rename remaining columns
```{r}
brain_names <- brain_names %>% 
rename(FS_LCort_GM_Vol = FS_L_CortVol_Vol,
FS_RCort_GM_Vol = FS_R_CortVol_Vol,
FS_TotCort_GM_Vol = FS_CortVol_Vol,   
FS_L_LatVent_Vol = FS_L_Lateral_Vent_Vol,
FS_L_InfLatVent_Vol = FS_L_Inf_Lat_Vent_Vol, 
FS_L_Cerebellum_WM_Vol = FS_L_Cerebellum_White_Matter_Vol,
FS_L_ThalamusProper_Vol = FS_L_Thalamus_Vol,
FS_3rdVent_Vol = FS_3rd_Vent_Vol, 
FS_4thVent_Vol = FS_4th_Vent_Vol,
FS_BrainStem_Vol = FS_Brain_Stem_Vol,
FS_L_Hippo_Vol = FS_L_Hippocampus_Vol,
FS_L_AccumbensArea_Vol = FS_L_Accumbens_Area_Vol,
FS_L_VentDC_Vol = FS_L_VentralDC_Vol,
FS_R_LatVent_Vol = FS_R_Lateral_Vent_Vol,
FS_R_InfLatVent_Vol = FS_R_Inf_Lat_Vent_Vol,
FS_R_Cerebellum_WM_Vol = FS_R_Cerebellum_White_Matter_Vol,
FS_R_ThalamusProper_Vol = FS_R_Thalamus_Vol,
FS_R_Hippo_Vol = FS_R_Hippocampus_Vol,
FS_R_AccumbensArea_Vol = FS_R_Accumbens_Area_Vol,
FS_R_VentDC_Vol = FS_R_VentralDC_Vol,
FS_CC_MidPosterior_Vol = FS_CC_Mid_Posterior_Vol,
FS_CC_MidAnterior_Vol = FS_CC_Mid_Anterior_Vol,
FS_InterCranial_Vol = ETIV.x.x,
FS_SupraTentorial_Vol = FS_SupraTentorialVol_Vol,
FS_SubCort_GM_Vol = FS_SubCortGrayVol_Vol)
```

remaining data features without corresponding model features:

FS_L_WhiteSurfArea_Area
BrainSegVolNotVent.x.x
FS_R_WhiteSurfArea_Area
BrainSegVolNotVent.y.x
ETIV.y.x
BrainSegVolNotVent.x.y
ETIV.x.y
BrainSegVolNotVent.y.y
ETIV.y.y
FS_L_vessel_Vol
FS_L_choroid_plexus_Vol
FS_R_vessel_Vol
FS_R_choroid_plexus_Vol
FS_5th_Vent_Vol
FS_WM_hypointensities_Vol
FS_L_WM_hypointensities_Vol
FS_R_WM_hypointensities_Vol
FS_Non_WM_hypointensities_Vol
FS_L_non_WM_hypointensities_Vol
FS_R_non_WM_hypointensities_Vol
FS_Optic_Chiasm_Vol
FS_BrainSegVol_Vol
FS_BrainSegVolNotVent_Vol
FS_LhCerebralWhiteMatterVol_Vol
FS_RhCerebralWhiteMatterVol_Vol
FS_CerebralWhiteMatterVol_Vol
FS_TotalGrayVol_Vol
FS_SupraTentorialVolNotVent_Vol
FS_MaskVol_Vol
FS_BrainSegVol_to_eTIV_Vol
FS_MaskVol_to_eTIV_Vol
FS_LhSurfaceHoles_Vol
FS_RhSurfaceHoles_Vol
FS_SurfaceHoles_Vol
FS_EstimatedTotalIntraCranialVol_Vol

cleaning
```{r}
#drop columns that aren't in model
common_cols <- intersect(colnames(training_sample_baseline), colnames(brain_names))

#test to make sure we don't have extras
col_test <- training_sample_baseline %>%
      select(!common_cols)

brain_subset <- brain_names %>% 
  select(Subid, common_cols)

```